### 2020-08-20 R/IPM practice.Rmd
initialize IPM practice Rmd

---
title: "Heliconia IPM Practice"
author: "Eric R. Scott"
date: "2020-08-20"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(skimr)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

Practice constructing an IPM using (a subset of) the [Heliconia demographic data](https://github.com/embruna/HeliconiaDataPaper/tree/master/data_clean).

# Load Data

```{r data, echo=TRUE}
demog <- read_csv(here("analysis", "data", "raw_data", "Ha_survey_with_Zombies.csv"), col_names = TRUE,
               cols(plot = col_character(),
                    bdffp_reserve_no = col_character(),
                    shts = col_integer(),
                    year = col_integer(),
                    infl = col_integer()))

unique(demog$ranch)
unique(demog$code.notes)
unique(demog$code2)
unique(demog$size)

skim(demog)
```

## Data Dictionary

`demog`: Demography data for *Heliconia acuminatata* experiment at BDFFP

- `plot` (character):
- `size` (character): fragment size
- `ranch` (character): "Dimona", "Esteio-Colosso", or "PortoAlegre".  The three ranches that make up BDFFP.
- `bdffp_reserve_no` (character): Official BDFFP reserve designation
- `tag_number` (numeric): plant tag# within plot (not unique)
- `row` (character): spatial location in plot
- `column` (numeric): spatial location in plot
- `year` (numeric): year of survey
- `ht` (numeric): plant height.  Plants not straightened before measurement.
- `shts` (numeric): number of shoots
- `infl` (numeric): number of infloresences
- `code.notes` (character): 
    - "plant missing" = plant was not found in that survey.  May not be dead, may be found in subsequent surveys.
    - "dead" = plant lost all above-ground tissue.  May resprout in a subsequent year.
    - "ULY" = 
    - "sdlg" = seedling, defined as a year 1 plant.
    - "tag missing" = ID tag labeling the plant was missing.  Plants were given new tags and researchers did their best to match up new tag and old tag. **(that column isn't in this version though)**
- `code2` (character): "initial.tag.yr" = marks the first year plant appeard in a survey.  Not necessarily a seedling.

# Analysis