### 2020-09-01 R/Workflow for yearly lambda.Rmd
Notes on how to programatically get lambdas for every transition year with a simple matrix model.

---
title: "R Notebook"
output: html_notebook
---

Use `popbio` to make a projection matrix

```{r}
library(tidyverse)
library(popbio)
```

One year at a time.  Good. That's what I want

```{r}
trans01 <- subset(merge(test.census, test.census, by = "plant", sort =FALSE),
                    year.x==2001 & year.y==2002 )
test.census %>% filter(year == 2002)
```

Add individual fertilities using "anonymous reproduction"  based on the proportional reproductive outputs of flowering plants and the total number of seedlings at the end of the projection interval

```{r}
# trans01$seedferts <- trans01$fruits.x / sum(trans01$fruits.x) * 5 #5 = number of seedlings in 2002.
trans01 <-
  trans01 %>% 
  mutate(seedferts = fruits.x / sum(fruits.x) * 5)

trans01
```
```{r}
stages <- c("seedling", "vegetative", "reproductive")
```

One way to specify columns:

```{r}
projection.matrix(trans01, stage.x, stage.y, seedferts, stages)
```

BEST to use column default (fertility column (seedling) now matches stage class name) AND order stages in dataframe

```{r}
# names(trans01)[c(3, 6, 8)] <- c("stage", "fate", "seedling")
# trans01$stage <- ordered(trans01$stage, stages)
trans01 <-
  rename(trans01, stage = stage.x, fate = stage.y, seedling = seedferts) %>% 
  mutate(stage = ordered(stage, c("seedling", "vegetative", "reproductive")))

trans01
```
Names in `stage` and `fate` have to match.  `stage` is ordinal. `seedling` column matches first stage to indicate fertility rates.

```{r}
projection.matrix(trans01)
```

# Try for every year

```{r}
census_wide <-
  test.census %>% 
  pivot_wider(names_from = year, values_from = c(stage, fruits))
```

```{r}
census_wide %>% 
  #need fertility columns for each year
  # mutate(fert_2001 = fruits_2001 / sum(fruits_2001, na.rm = TRUE) * sum(str_detect(stage_2002, "seedling"), na.rm = TRUE))
  mutate(across(starts_with("fruits"), ~ . / sum(., na.rm = TRUE), .names = "{col}.tot"))
```
```{r}
#complete data frame with all plant IDs.  Necessary for lags to work right.
full.census <-
  left_join(
    test.census %>% 
      expand(plant = 1:18, year) %>% 
      arrange(year),
    test.census
  )

full.census <-
  full.census %>%
  # generate transitions
  group_by(plant) %>% 
  mutate(transition = paste0(year, "-", lead(year))) %>% 
  mutate(fate = lead(stage)) %>% 
  # calculate anonymous reproduction
  group_by(year) %>% 
  mutate(num_seedling = sum(str_detect(fate, "seedling"), na.rm = TRUE)) %>% 
  mutate(seedling = fruits / sum(fruits, na.rm = TRUE) * num_seedling) %>% 
  # remove plants that started off a transition dead
  filter(stage != "dead", !is.na(fate)) %>% 
  mutate(stage = ordered(stage, c("seedling", "vegetative", "reproductive")))
```

```{r}
testyr <- full.census %>% filter(year == 2001)
projection.matrix(as.data.frame(testyr), stage = stage, fate = fate, fertility = seedling)
projection.matrix(trans01, stage = stage, fate = fate, fertility = seedling)
projection.matrix(full.census %>% as.data.frame())
```

Get lambda for every transition

```{r}
full.census %>% 
  as.data.frame() %>%  
  group_by(transition) %>% 
  nest() %>% 
  mutate(K = map(data,
                 ~projection.matrix(as.data.frame(.x))
  )) %>% #projection.matrix() doesn't work with tibbles for some reason
  mutate(lambda = map_dbl(K, ~ lambda(.x)))
```