### 2020-09-03 notes/IPM practice.Rmd

---
title: "Heliconia IPM Practice"
author: "Eric R. Scott"
date: "2020-08-20"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(skimr)
library(performance)
library(lme4)
library(bbmle)
library(broom)
library(broom.mixed)
library(popbio)
library(janitor)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("slice", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# TODO:

- I feel like there is maybe a more sophisticated method of dealing with "zombie" plants.  Take another look at Elizabeth's lectures or whats-her-name's paper on terrestrial orchids.
- Not sure how to incorporate overdispersion (significant random effect of `tag_number`)

# Purpose

Practice constructing an IPM using (a subset of) the [Heliconia demographic data](https://github.com/embruna/HeliconiaDataPaper/tree/master/data_clean).

# Load Data

```{r data, echo=TRUE}
demog <- read_csv(here("analysis", "data", "raw_data", "Ha_survey_with_Zombies.csv"), col_names = TRUE,
               cols(plot = col_character(),
                    bdffp_reserve_no = col_character(),
                    shts = col_integer(),
                    year = col_integer(),
                    infl = col_integer(),
                    tag_number = col_character(),
                    HA_ID_Number = col_character())) %>% 
  clean_names()

unique(demog$ranch)
unique(demog$code_notes)
unique(demog$code2)
unique(demog$habitat)

skim(demog)
```

## Data Dictionary

`demog`: Demography data for *Heliconia acuminatata* experiment at BDFFP

- `plot` (character):
- `habitat` (character): fragment size
- `ranch` (character): "DIM" = Dimona, "EST" = Esteio-Colosso, or "PAL" = PortoAlegre.  The three ranches that make up BDFFP.
- `bdffp_reserve_no` (character): Official BDFFP reserve designation
- `ha_id_number` (numeric): A unique ID number for each plant after resolving duplicate tags and such.
- `tag_number` (numeric): plant tag# within plot (not unique)
- `row` (character): spatial location in plot
- `column` (numeric): spatial location in plot
- `year` (numeric): year of survey
- `ht` (numeric): plant height.  Plants not straightened before measurement.
- `shts` (numeric): number of shoots
- `infl` (numeric): number of inflorescences
- `code.notes` (character): 
    - "plant missing" = plant was not found in that survey.  May not be dead, may be found in subsequent surveys.
    - "dead" = plant lost all above-ground tissue.  May resprout in a subsequent year.
    - "ULY" = 
    - "sdlg" = seedling, defined as a year 1 plant.
    - "tag missing" = ID tag labeling the plant was missing.  Plants were given new tags and researchers did their best to match up new tag and old tag.
- `code2` (character): "initial.tag.yr" = marks the first year plant appeared in a survey.  Not necessarily a seedling.

# Data Wrangling

Take subset for messing around

```{r}
km41 <- demog %>% filter(bdffp_reserve_no == "1501") %>% select(-bdffp_reserve_no)
km41 %>% count(plot)
```

Narrow down to just one plot for now

```{r}
km41 <- km41 %>% filter(plot == "5750")
km41 %>% count(tag_number) %>% count(n)

km41 %>% count(tag_number) %>% ggplot() + geom_histogram(aes(n))
```
Plants have up to 12 years of data

## Survival and reproduction

I need to have a column for fate.  Made difficult by "zombie" plants.

```{r}
km41
```
When is a plant really dead?  When it's the last non-`NA` value for that particular tag number before the max date.

```{r}
km41 %>% group_by(ha_id_number) %>% summarize(max(year))
max(km41$year)
```

```{r}
km41_2 <- 
  km41 %>%
  group_by(ha_id_number) %>% 
  # Routine for figuring out last year of data collected
  mutate(x = cumsum(ht > 0 & !is.na(ht))) %>%
  # mutate(x = cumsum(lag(ht) > 0 & !is.na(lag(ht)))) %>%
  mutate(n = 1:n()) %>% 
  mutate(fate = if_else(
    n == which.max(x==max(x)),
    "died",
    NA_character_
  )) %>% 
  # remove rows after plant is dead
  filter(row_number() <= which(fate == "died")[1]) %>% 
  # remove "died" if last year of data (2009)
  ungroup() %>% 
  mutate(fate = if_else(year == max(year), NA_character_, fate)) %>% 
  select(-n, -x)
# km41_2 %>% filter(fate == "died") %>% count(year)
```
```{r}
km41_2 %>% count(fate)
```
362 plants died

```{r}
km41_2 <- km41_2 %>% mutate(surv = ifelse(is.na(fate), 1, 0))
```

```{r}
km41_2 <- km41_2 %>% mutate(flowered = infl > 0)
```

## Just flowering individuals

```{r}
km41_flowered <- km41 %>% filter(infl > 0)
```

## Lagged height for growth

```{r}
km41_2 <-
  km41_2 %>% 
  group_by(ha_id_number) %>% 
  mutate(ht_prev = lag(ht))
```

# IPM

Gotta look this up, but assuming that reproduction happens before survival

## Kernel

In words:

The number of individuals in a size class in year t+1 is a function of

$$
K(z^\prime,z) = p_b(z)b(z)p_rc_0(z^\prime) + s(z)G(z, z^\prime)
$$
Reporduction:

- $p_b(z)$: probability of flowering as a function of size
- $b(z)$: number of inflorescences as a function of size. 
- $p_r$ mean # of seedlings produced per inflorescence.  This varies by year, not sure how to incorporate environmental stochasticity.
- $c_0(z^\prime)$ size distribution of new recruits

Growth & survival:

- $s(z)$ is survival probability as a function of size
- $G(z, z^\prime)$ is growth as a function of size.

## Reproduction

In this particular dataset the only information on reproduction is number of inflorescences per plant and total number of seedlings in a year.  However, Bruna et al. 2004 has some additional data on reproduction:

- each inflorescence has 20--25 flowers
- proportion of flowers developing into fruits = 0.30±0.24 (range = 0--0.88)
- Number of seeds per plant = 4.74±6.10 (range = 0--27)
- Number of sees per fruit =  1, 2, or 3 with equal probability.
- Number of seeds per flower = 0.27±0.33 (range = 0--1)

But the last step, seedlings per seed, isn't there, so I'm not sure if this info is relevant for this IPM, especially if none of the above vital rates is expected to differ by plant size.

### Probability of flowering

#### Model

Start with most complex: quadratic response, environmental stochasticity, and random variation among plants.

```{r}
meo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
```
Check random effects

```{r}
mo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|ha_id_number), family = binomial, data = km41_2)
me_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
m_pflwr2  <-   glm(flowered ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_pflwr2, me_pflwr2, mo_pflwr2, m_pflwr2)
```
Keep both random effects.

Check quadratic

```{r}
meo_pflwr <- glmer(flowered ~ ht + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
AICtab(meo_pflwr, meo_pflwr2)
```

Keep quadratic

```{r}
augment(meo_pflwr2, type.predict = "response") %>% 
  # filter(tag_number %in% c(1, 101)) %>%
  ggplot(aes(x = ht, y = .fitted)) +
  geom_line(alpha = 0.2, aes(group = year)) +
  stat_summary_bin(geom = "point", fun = mean, aes(y = as.numeric(flowered)), bins = 80, alpha = 0.5)
```

Some years had super high survival, others very low.  Variation among plants.

#### Function

```{r}
fixef(meo_pflwr2)
```

$$
p(flower) = inv.logit(-7.423 + 0.181ht - 0.0008ht^2)
$$

Not sure how to incorporate random effects

```{r}
p_bz <- function(z) {
  coefs <- fixef(meo_pflwr2)
  linear.p <- coefs[1] + coefs[2]*z + coefs[3]*z^2
  p <- 1/(1+exp(-linear.p))
  return(as.numeric(p))
}
# predict(mm_pflwr2, newdata = data.frame(ht = 20, tag_number = "new"), allow.new.levels = TRUE) %>% plogis() ==
# p_bz(20)
```

### Number of infloresences

#### Model

```{r}
meo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = poisson, data = km41_flowered)
```
```{r}
me_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|year), family = poisson, data = km41_flowered)
mo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|ha_id_number), family = poisson, data = km41_flowered)
m_flwrs2 <- glm(infl ~ ht + I(ht^2), family = poisson, data = km41_flowered)
AICtab(meo_flwrs2, me_flwrs2, mo_flwrs2, m_flwrs2)
```

No random effects.

Check quadratic

```{r}
m_flwrs <- glm(infl ~ ht, family = poisson, data = km41_flowered)
AICtab(m_flwrs, m_flwrs2)
```

Go with simpler linear glm

```{r}
augment(m_flwrs, type.predict = "response") %>% 
  ggplot(aes(x = ht, y = .fitted)) +
  geom_line() +
  geom_jitter(aes(y = infl), alpha = 0.5, width = 0, height = 0.1)
```

#### Function

```{r}
b_z <- function(z) {
  coefs <- coef(m_flwrs)
  infl <- coefs[1] + coefs[2]*z
  return(as.numeric(infl))
}
```

### Probability of recruitment

#### Model

Estimated as number of recruits / number of inflorescences in year prev. year

```{r}
km41_rec_sum <- 
  km41 %>%
  group_by(year) %>% 
  summarize(recruits = sum(code_notes == "sdlg (1)", na.rm = TRUE),
            seeds_est = sum(lag(infl), na.rm = TRUE)) %>% 
  filter(year > 1998)
km41_rec_sum
```

```{r}
m <- glm(recruits ~ 1, offset = log(seeds_est), family = poisson, data = km41_rec_sum)
coef(m)
```

#### Function

```{r}
p_r <- as.numeric(exp(coef(m)))
p_r
```

So like 1.64 seedlings per inflorescence

### Size distribution of new recruits

#### Model

recruitment size

```{r}
km41_rec <- km41 %>% filter(code_notes == "sdlg (1)")
m_rec <- lm(ht ~ 1, data = km41_rec)
```

```{r}
coef(m_rec)
summary(m_rec)$sigma
```

#### Function

```{r}
c_0z1 <- function(z1) {
  mu <- coef(m_rec)[1]
  sig <- summary(m_rec)$sigma
  p.deRecr <- dnorm(z1, mean = mu, sd = sig)
  return(p.deRecr)
}
# c_0z1(20)
```

## Growth

### Model

Try modeling growth as height_t+1 ~ height_t.  Needs a random effect of tag_number, I think?
```{r}
meo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|ha_id_number) + (1|year), data = km41_2)
```
Test random effects

```{r}
me_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|year), data = km41_2)
mo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|ha_id_number), data = km41_2)
m_grow2 <- lm(ht ~ ht_prev + I(ht_prev^2), data = km41_2)
AICtab(meo_grow2, me_grow2, mo_grow2, m_grow2)
```

Environmental stochastsicity, but not differences among plants

check quadratic

```{r}
me_grow <- lmer(ht ~ ht_prev + (1|year), data = km41_2)
AICtab(me_grow, me_grow2)
```

Quadratic is better.

```{r}
augment(me_grow2) %>% 
ggplot(aes(x = ht_prev, y = ht)) + 
  geom_point(alpha = 0.2) +
  geom_line(aes(y = .fitted, group = year), color = "blue", alpha = 0.2)
```
### Function

```{r}
fixef(me_grow2)
```

$$
ht_{t+1} = 2.143 + 1.056ht - 0.002ht^2
$$

```{r}
summary(me_grow2)$sigma
summary(me_grow2)$varcor
```

Not sure how to take random effect into account

```{r}
G_z1z <- function(z1, z) {
  coefs <- fixef(me_grow2)
  sig <- summary(me_grow2)$sigma #or shoudl this incorporate mixed effects somehow?

  mu <- coefs[1] + coefs[2]*z + coefs[3]*z^2
  p.den.grow <- dnorm(z1, mean = mu, sd = sig) #pdf of new size for current size
  return(p.den.grow)
}
```

## Survival

### Model
```{r}
meo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
```
```{r}
me_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
mo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|ha_id_number), family = binomial, data = km41_2)
m_surv2 <- glm(surv ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_surv2, me_surv2, mo_surv2, m_surv2)
```

Only environmental stochasticity.

Check quadratic

```{r}
me_surv <- glmer(surv ~ ht + (1|year), family = binomial, data = km41_2)

AICtab(me_surv, me_surv2)
```

Quadratic is better

```{r}
augment(me_surv2, type.predict = "response") %>% 
  ggplot(aes(x = ht, y = .fitted)) +
  geom_line(aes(group = year), color = "blue", alpha = 0.5) +
  stat_summary_bin(geom = "point", fun = mean, aes(y = surv), bins = 80, alpha = 0.5)
```

Survival is better in some years than others, especially for very small plants

### Function

```{r}
fixef(me_surv2)
```

$$
p(survive) = inv.logit(0.920 + 0.142ht - 0.0009ht^2)
$$

```{r}
s_z <- function(z) {
  coefs <- fixef(me_surv2)
  linear.p <- coefs[1] + coefs[2]*z + coefs[3]*z^2
  p <- 1/(1+exp(-linear.p))
  return(as.numeric(p))
}
```

# Fecundity kernel

p_b(z)b(z)p_rc_0(z^\prime)

```{r}
F_z1z <- function(z1, z) {
  return(p_bz(z) * b_z(z) * p_r * c_0z1(z1))
}
```

# Survival kernel

s(z)G(z, z^\prime)

```{r}
P_z1z <- function(z1, z) {
  return(s_z(z) * G_z1z(z, z1))
}
```

# Run IPM

## Decide meshpoints

```{r}
range(km41$ht[km41$ht>0], na.rm = TRUE)
```

So I'll go from 0.4 to 200??

```{r}
hist(km41$ht[km41$ht>0])
```
## Projection matrix

```{r}
mk_K <- function(m, L, U) {
	# mesh points 
	h <- (U - L)/m
	meshpts <- L + ((1:m) - 1/2) * h
	P <- h * (outer(meshpts, meshpts, P_z1z))
	F <- h * (outer(meshpts, meshpts, F_z1z))
	K <- P + F
	return(list(K = K, meshpts = meshpts, P = P, F = F))
}
```

```{r}
out <- mk_K(100, 0.4, 200)
```

# Results

```{r}
lambda(out$K)
plot(out$meshpts, stable.stage(out$K), type = "l")
```

```{r}
hist(km41$ht)
```
Kinda matches?

```{r}
sens <- sensitivity(out$K)

rownames(sens) <- out$meshpts
colnames(sens) <- out$meshpts
matplot2(sens, xlab = "Height")
```

Can't remember how to interpret this

```{r}
elas <- elasticity(out$K)
rownames(elas) <- colnames(elas) <- out$meshpts
matplot2(elas, xlab = "Height")
```