### 2020-09-23 R/competition wrangling.Rmd
filtered out plants near edges

---
title: "Competition"
author: "Eric R. Scott"
date: "2020-09-09"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(glue)
library(car)
library(skimr)
library(furrr)
library(janitor)

source(here("R", "utils.R"))

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

# TODO:

- filter out focal plants where donut goes out of plot. (filter(distance_to_edge > outer_circle_radius))
- Using all years is not estimating density accurately. Focal plants are not moving, but in other years there are neighbors without known coordinates. Solution: only use the three transitions where we know the coords of every plant?  Or assume births and deaths don't change density significantly?

# Load Data

```{r data, echo=TRUE}
# demographic data
ha <- read_rds(here("analysis", "data", "derived_data", "ha_survey.rds"))
# data from previous study: Uriarte M, Bruna EM, Rubim P, et al (2010) Effects
# of forest fragmentation on the seedling recruitment of a tropical herb:
# assessing seed vs. safe-site limitation. Ecology 91:1317–1328.
# https://doi.org/10.1890/09-0785.1

coords <-
  read.table(here("analysis", "data", "raw_data", "adults0607_light.txt")) %>%
  as_tibble() %>%
  clean_names()
```

`ha` contains x and y coords for each plant from 2009 *within each grid cell*.  `coords` contains x and y coordinates for each plant in 2006 and 2007 relative to the entire plot.  I need to join these data sources and offset the coordinates in the `ha` dataset correctly.  I'm going to remove `year` from the `coords` dataset and assume that plants aren't moving.

# Join data

Getting things to match up:

```{r}
coords2 <- 
  coords %>% 
  select(plot = site, tag_number, row, column, plot_x, plot_y, x, y) %>% 
  mutate(plot = case_when(
    plot == "Cabo Frio" ~ "CaboFrio-CF",
    plot == "Dimona CF" ~ "Dimona-CF",
    plot == "Florestal" ~ "Florestal-CF",
    plot == "PA-CF" ~ "PortoAlegre-CF",
    TRUE ~ plot
  )) 
```

Summarize so there is only one set of coordinates per plant. Averaging as I'm assuming any difference in coordinates between '06 and '07 is not because the plant moved.

```{r}
coords3 <-
  coords2 %>%
  group_by(plot, tag_number) %>% 
  # rowwise() %>% 
  summarize(across(where(is.character), first),
            across(where(is.numeric), mean)) %>% 
  rename(x_07 = x,
         y_07 = y)
```

```{r}
ha_coords <- left_join(ha, coords3, by = c("plot", "row", "column", "tag_number"))
```
# Tidy

Now, use the plot_x and plot_y variables to correctly offset the x and y coords from 2009.

```{r}
ha_coords2 <- 
  ha_coords %>% 
  mutate(x_adj = x_09 + ((plot_x - 1) * 10),
         y_adj = y_09 + ((plot_y - 1) * 10),
         .after = y_09)
```

Now, I'll just average across the three years and assume that any differences are measurement error and not because plants moved.

```{r}
ha_coords3 <-
  ha_coords2 %>% 
  rowwise() %>% 
  mutate(x = mean(c(x_adj, x_07), na.rm = TRUE),
         y = mean(c(y_adj, y_07), na.rm = TRUE)) %>% 
  #replace NaN's generated when mean of only NAs is taken
  mutate(across(c(x, y), ~ifelse(is.nan(.), NA, .))) %>% 
  #remove intermediate coordinates
  select(ranch, bdffp_reserve_no, plot, habitat,
         ha_id_number, tag_number, row, column,
         x, y, year, ht, ht_next, shts, shts_next,
         infl, surv, code_notes, code2)
```

# Filter

To model the effect of density, I'm only going to use the subset of plants that I have coordinates for.

**NOTE:** There is a **problem** with this.  By using all the years and filtering the data to just include plants with coordinates from 2006, 2007, and 2009, I am not estimating density accurately.
```{r}
ha_xy <-
  ha_coords3 %>% 
  filter(!is.na(x), !is.na(y)) %>% 
  ungroup()
```

# Calculate density

Now the challenge is to calculate the density of surrounding shoots for every focal plant at different distances.

```{r}
# function to calculate euclidean distance between two points
eu_dist <- function(p1, p2) {
  sqrt(sum((p1 - p2)^2))
}
# x = c(0,0)
# y = c(3,3)
# eu_dist(x, y)
```

## Proof of concept

Start with single plot

```{r}
y1p1 <- ha_xy %>% 
  filter(year == first(year),
         plot == first(plot))
y1p1
```

```{r}
# df <- y1p1
# L = 0
# U = 10
# by = 0.5
add_density_cols <- function(df, L, U, by, na_edges = TRUE) {

  eu_dist <- function(p1, p2) {
    sqrt(sum((p1 - p2)^2))
  }

  d <- seq(L, U, by = by)

  df_out <- 
    future_map_dfr(1:nrow(df), function(r) { #loop through rows/plants
      #get coordinates for that plant
      focal_plant <- t(df[r, c("x", "y")])
      # focal_plant <- c(0,0)

      if (na_edges) {
        #don't do calculations for rings that go outside the plot
        edge_dist <- min(c(focal_plant, c(100, 50) - focal_plant))
        d2 <- d[which(d < edge_dist)]
      } else {
        d2 <- d
      }

      #calculate distance to every other plant
      focal_df <-
        df %>%
        rowwise() %>%
        mutate(dist = eu_dist(focal_plant, c(x, y))) %>% 
        ungroup()

      # If plant is near edge, return NAs

      #get total number of shoots at every ring
      if (length(d2) > 1) {
        future_map2_dfc(lag(d2)[-1], d2[-1],
                        ~ transmute(
                          focal_df,
                          !!glue("shts_{.x}_{.y}") :=
                            shts * (dist > .x & dist <= .y)
                        )
        ) %>% 
          #excludes focal plant with dist == 0
          summarize(across(everything(), ~sum(., na.rm = TRUE)), .groups = "drop")
      } else {
        # if all the rings go outside the plot, then return an empty row
        tibble_row()
      }
    })

  df_out <- bind_cols(df, df_out) #join to original data
  return(df_out)
}

add_density_cols(y1p1, L = 0, U = 10, by = 5)

```

If I want this to be in units of shoots/m^2, then I need to calculate the area of the donut:

$$
A = \pi(U^2 - L^2)
$$
But also, like, truncate that donut if it passes outside of the plot.  I don't know how to do that.

```{r}
plan(multiprocess)
```

## Apply to entire dataset

```{r}
ha_xy2 <-
  ha_xy %>% 
  group_by(plot, year) %>% 
  group_split() %>% 
  # head() %>% #test with 6 df to see how long it will take.
  future_map(~{
      add_density_cols(.x, L = 0, U = 10, by = 0.5) 
    }) %>% 
  bind_rows()
beepr::beep(4)
```

Aggregate some of the distances for simpler modeling.

```{r}
ha_xy3 <-
  ha_xy2 %>% 
  rowwise() %>%
  mutate(shts_0_5 = sum(c_across(shts_0_0.5:shts_4.5_5)),
         shts_5_10 = sum(c_across(shts_5_5.5:shts_9.5_10)))

ha_density <- ha_xy3
```

Yay! It worked! And it didn't even run that long!

# Write to file

```{r}
write_rds(ha_density, here("analysis", "data", "derived_data", "ha_density.rds"))
```