### 2020-10-28 analysis/paper/appendix_density_dependence.Rmd
revised row and column locations and re-ran density dependence analysis.

---
title: "Density dependence"
author: "Eric R. Scott"
date: '2020-09-13'
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  include = FALSE,
  echo = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300)
library(tidyverse)
library(here)
library(conflicted)
library(bbmle)
library(dlnm)
library(mgcv)
library(broom)
library(broom.mixed)
library(lme4)
library(car)

source(here("R", "crossbasis.R"))
source(here("R", "utils.R"))

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# TODO

-   figure captions

-   revise

-   equations for models

-   Explain why heatmap is missing a chunk (search literature for other 2d gam papers that use `too.far` argument?)

-   Update stats in writing

# Methods

## Data collection

The exact coordinates (to the nearest 0.1m) of each plant relative to the plot were recorded for a subset of plots in three years. In 2006 and 2007, locations of plants were recorded to the nearest meter as part of a study to assess the effects of safe-site limitation on recruitment [@uriarte2010]. In 2009, plants were mapped to .... <!--# rubim study -->For each plot and year, we calculated the density of conspecific shoots in annuli at varying distances from focal plants, as in Teller et al. [-@teller2016]. Density was calculated for 5 annuli ranging from an inner radius of 0m and an outer radius of 0.5m (i.e. a circle of radius 0.5m) to an inner radius of 4.5m and and outer radius of 5m. If an annulus intersected with the edge of the plot, the density for that particular plant and annulus was considered as a missing value. Conspecific density was estimated as the size of H. acuminata shoots per m2 in each annulus. <!--# Add analyses using height and shoots + height -->Number of shoots was used as a measure of plant size rather than height as it is less error prone to measure, is not affected by plants being knocked over, and is strongly correlated with plant height and leaf area in H. acuminata [@bruna2002; @brunaEffectsForestFragmentation2002].

We used the three transition years for which we had density data to test for effects of conspecific density on growth and survival of *H. acuminata*. We were unable to test for the effects of density dependence on probability of flowering, as surveys of reproduction were not conducted in these years <!--# or perhaps no plants flowered? -->.

```{r data}
ha <- read_rds(here("analysis", "data", "derived_data", "ha_density.rds"))
ha$year %>% range()
ha$code_notes %>% unique()
```

```{r tidy}
ha <-
  ha %>%
  filter(code_notes != "sdlg (1)" | is.na(code_notes)) %>% #remove seedlings
  arrange(ha_id_number, year) %>% 
  #scale year, and convert to factor since I'll be using as random effect
  mutate(year_post = as_factor(year - min(year)),
         log_shts = log(shts),
         log_shts_next = log(shts_next),
         # log_ht = log(ht),
         # log_ht_next = log(ht_next),
         habitat = as.factor(habitat),
         ha_id_number = as.factor(ha_id_number),
         plot = as.factor(plot))
```

## Statistical Modeling

We modeled density dependence using a penalized distributed lag non-linear model (DLNM) approach with the `dlnm` package in R [@gasparriniPenalizedFrameworkDistributed2017; @gasparriniDistributedLagLinear2011; @rcoreteamLanguageEnvironmentStatistical2020]. This approach is similar to that of Teller et al. [-@teller2016], in that it makes no assumptions about the shape of the competition kernel. Instead, it fits a two-dimensional spline allowing for a non-linear response to the intensity of conspecific density as well as the distance of conspecifics. We implemented this in a generalized additive model (GAM) framework taking advantage of the penalization features of the `mgcv` package [@woodGeneralizedAdditiveModels2017].

```{r}
Q <-
  ha %>%
  select(shts_0_0.5:shts_4.5_5) %>%
  as.matrix()

L <-
  matrix(seq(0.25, 4.75, by = 0.5), #use midpoint of r1 and r2
         nrow = nrow(Q),
         ncol = ncol(Q),
         byrow = TRUE)
```

```{r}
g_growth_scat <-
  gam(shts_next ~ 
        # s(ha_id_number, bs = "re") + #this doesn't seem to converge
        s(plot, bs = "re") +
        # s(log_shts, bs = "cr", by = habitat) + #prelim modeling shows this is a linear relationship (edf ~ 1)
        log_shts *
        habitat +
        s(Q, L,
          bs = "cb",
          # k = c(3,3),
          xt = list(bs = c("cr"))),
      family = scat(link = "log"),
      data = ha,
      method = "REML")
```

Growth was modeled as: (still not sure how to do this. Summation is from Gasparini et al, but also written as simply as $s(\textbf{q}_{i,t})$

$$
size_{t+1} = \sum_{d=d_0}^D f \cdot w(x_{i,d}, d)
$$

Where $f \cdot w(x, d)$ is a two dimensional crossbasis smooth composed of two marginal functions: $f(x)$, the effect of conspecific density, $x$ , on growth, and the additional distance response function $w(d)$ that models the effect of density at different annuli, $d$, equal to the midpoint of the inner and outer radius of the 5 annuli for which density was calculated. Each dimension was modeled with 10 knots. With a penalized spline, the number of knots is not very important as ... <!--# finish thought and cite (maybe Noam Ross) --> Habitat is a parametric fixed effect with two levels (1-ha and continuous forest), $s(log(shts_t))$ is a penalized cubic spline describing the growth from year t to year t + 1, and a random effect of plot, nested within habitat. We did not include a random effect of plant ID, as the model did not converge. Although the response, shoot number in year t+1, is a discrete variable, a scaled t distribution with a log-link was used as it was a better fit than a poisson distribution ($\Delta$AIC = 810.5) and a log-normal distribution ($\Delta$AIC = 294).

```{r eval=FALSE, include=FALSE}
g_growth_pois <- update(g_growth_scat, family = poisson(link = "log"))
g_growth_norm <- update(g_growth_scat, family = gaussian(link = "log")) #residuals are leptokurtic with gaussian
AICtab(g_growth_norm, g_growth_pois, g_growth_scat)
```

Survival was modeled using the same predictors as for growth, but with a binomial family error distribution.

```{r}
g_surv <-
  gam(surv ~
        s(plot, bs = "re") +
        s(log_shts, bs = "cr", by = habitat) +
        habitat +
        s(Q, L,
          bs = "cb",
          xt = list(bs = "cr")),
      family = binomial,
      data = ha,
      method = "REML")
# plot(g_surv)
```

Because the DLNM approach does not allow for modeling an interaction between conspecific density and habitat type, we used the outputs of these cross-basis smooths to fit generalized linear mixed-effects models with interaction terms. By inspecting the marginal effects plots from the cross-basis smooths, we identified a range of distances over which conspecific density had the greatest effect and used density over that distance as a predictor in the GLMMs. <!--# revise and also talk more about what interactions and random effects included -->

# Results

## Growth

```{r}
# g_growth2 <- update(g_growth, ~-s(plot, bs = "re"))
# AICtab(g_growth, g_growth2)
# g_growth3 <- update(g_growth, ~-s(log_shts) + log_shts)
# AICtab(g_growth, g_growth2, g_growth3)
# Best model:
anova(g_growth_scat)
# plot(g_growth_scat)
```

```{r growth-plot, include=TRUE, fig.cap="this is a figure"}
Q_new <- 
  g_growth_scat$model %>% 
  pull(Q) %>%
  as_tibble() %>% 
  summarize(across(everything(), mean)) %>%
  as.matrix()
L_new <- L[1,, drop = FALSE]

newdata <- tibble(log_shts = seq(0, 2.565, length.out = 100), habitat = "1-ha", plot = "new", Q = Q_new, L = L_new)

fitted <-
  augment(
    g_growth_scat,
    newdata = newdata,
    se_fit = TRUE,
    conf.level = 0.95
  )
ggplot(fitted, aes(x = exp(log_shts), y = exp(.fitted))) +
  geom_jitter(data = ha %>%
                filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))),
              aes(x = shts, y = shts_next),
              alpha = 0.2) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = exp(.fitted - .se.fit*1.96), ymax = exp(.fitted + .se.fit*1.96)), alpha = 0.4) +
  labs(x = "#shoots (t)", y = "#shoots (t + 1)") +
  theme_bw()
```

Number of shoots in the current year had a significant effect on size in year t + 1 ($\chi^2$ = 796.8, df = 1, P \< 0.0001) (Figure \@ref(fig:growth-plot)). There was no significant effect of habitat or the interaction between habitat and size on size in year t + 1. The cross-basis smooth describing the effects of conspecific density on growth was significant ($\chi^2$ = 23.2, edf = 6.4, P = 0.004).

```{r}
pred_growth <- cb_margeff(Q, L, g_growth_scat)
```

```{r growth-cb, include=TRUE, fig.cap="this is a figure"}
pred_growth %>%
  filter(min_dist < 0.1) %>%
  ggplot(aes(x = x, y = lag, z = exp(fitted), fill = exp(fitted))) +
  geom_raster(interpolate = TRUE) +
  # geom_contour(color = "grey50", size = 0.4) + 
  scale_y_continuous("distance (m)", expand = c(0,0)) +
  scale_x_continuous("conspecific density (shoots/m^2)", expand = c(0,0)) +
  scale_fill_viridis_c("#shoots (t+1)", option = "A") +
  theme_bw()
```

Plants had a greater number of shoots in year t+1 as conspecific density increased, all else being average. The effects of conspecific density were strongest around 0--1 meters from the focal plant (Figure \@ref(fig:growth-cb)).

```{r growth-slice, include=TRUE, fig.cap="this is a figure"}
pred_growth %>%
  filter(min_dist < 0.1) %>%
  filter(nearest(lag, 0.5)) %>% 
  ggplot(aes(x = x, y = exp(fitted), ymin = exp(fitted - 1.96*se.fit), ymax = exp(fitted + 1.96*se.fit))) +
  geom_jitter(data = ha %>%
                filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))),
              aes(x = shts_0_0.5, y = shts_next),
              alpha = 0.2,
              inherit.aes = FALSE) +
  geom_line(color = "blue") + 
  geom_ribbon(alpha = 0.4) +

  labs(title = "Near 0.5m from focal plant", 
       x = "conspecific density (shoots/m^2)",
       y = "#shoots (t+1)") +
  theme_bw()
```

In that annulus, conspecific density showed no effect (i.e. a flat line) below around 4 shoots m^-2^ (Figure \@ref(fig:growth-slice)). Only a few focal plants had a conspecific density greater than 4 shoots m^-2^, so the support for positive density dependence in this spline is driven strongly by only a few points. Additionally, it is important to note that p-values from splines in generalized additive model are approximate and likely too low [@woodGeneralizedAdditiveModels2017].

## Survival

```{r}
# plot_gam_check(g_surv)
anova(g_surv)
```

Plant size had a significant, nearly quadratic (on a logit scale), effect on survival in continuous forest ($\chi^2$ = 112.2, edf = 2.1, P \< 0.001), but a linear (on a logit scale) effect in 1 ha fragments ($\chi^2$ = 15.63, edf = 1, P \< 0.001), but their 95% confidence intervals overlap (Figure \@ref(fig:surv-plot)). However, there were no significant effects of habitat ($\chi^2$ = 0.002, df = 1, P = 0.964) or conspecific density ($\chi^2$ = 5.36, edf = 2, P = 0.0688) on survival.

```{r surv-plot, include=TRUE, fig.cap="this is a figure"}
Q_new <- 
  g_surv$model %>% 
  pull(Q) %>%
  as_tibble() %>% 
  summarize(across(everything(), mean)) %>%
  as.matrix()
L_new <- L[1,, drop = FALSE]

#TODO: make this have both 1-ha and CF

newdata <-
  tibble(
    log_shts = rep(seq(0, 2.565, length.out = 100), 2),
    habitat = rep(c("1-ha", "CF"), each = 100),
    plot = "new",
    Q = Q_new,
    L = L_new
  )

fitted <- augment(g_surv, newdata = newdata, se_fit = TRUE)

#for rug
rugdat <- 
  ha %>%
  filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))) %>% 
  mutate(survived = ifelse(surv == 1, shts, NA),
         died = ifelse(surv == 0, shts, NA))

ggplot(fitted, aes(x = exp(log_shts), y = plogis(.fitted), color = habitat)) +
  geom_line() +
  geom_ribbon(aes(
    ymin = plogis(.fitted - .se.fit * 1.96),
    ymax = plogis(.fitted + .se.fit * 1.96),
    color = NULL,
    fill = habitat
  ), alpha = 0.4)+
  geom_rug(data = rugdat,
           aes(x = survived, color = habitat),
           inherit.aes = FALSE,
           sides = "t",
           alpha = 0.2) +
  geom_rug(data = rugdat,
           aes(x = died, color = habitat),
           inherit.aes = FALSE,
           sides = "b",
           alpha = 0.2) +
  stat_summary_bin(geom = "point",
                   fun = mean,
                   aes(y = surv),
                   bins = 1000,
                   data = ha %>%
                     filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.)))) +
  labs(x = "#shoots (t)", y = "survival probability") +
  scale_x_continuous(n.breaks = 7) +
  theme_bw()
```

## Mixed Models Approach

Because the DLNM approach does not allow for interaction terms and did not converge with the inclusion of certain random effects, we repeated the analysis using a more standard generalized linear mixed effects model approach. Because the effects of conspecific density dropped off steeply beyond 3m from the focal plant (Figure \@ref(fig:growth-cb)), we re-calculated conspecific shoot density within a radius of 3m and used this as a predictor. We used AIC comparisons to test for significance of random effects, then interactions, then used a marginal hypothesis test to test for significance of main effects of log(shoot number), habitat, and conspecific density. Marginal hypothesis tests were conducted with the `Anova` function from the `car` package [@foxCompanionAppliedRegression2019].

```{r}
mnorm <-
  glmer(
    shts_next ~ 
      log_shts *
      habitat *
      shts_0_1 +
      (1|ha_id_number) +
      (1|plot) +
      (1|year_post),
    family = gaussian(link = "log"), #can't use scat() in glmer()
    data = ha
  )
# Random effects
mnorm2 <- update(mnorm, . ~ . - (1|ha_id_number))
AICtab(mnorm, mnorm2) #keep ID number
mnorm3 <- update(mnorm, .~. - (1|plot)) 
AICtab(mnorm, mnorm2, mnorm3) #remove plot
mnorm4 <- update(mnorm3, .~.-(1|year_post)) #remove year
AICtab(mnorm, mnorm2, mnorm3, mnorm4)

# Interactions
mnorm5 <- update(mnorm4, .~. - log_shts:habitat:shts_0_1)
AICtab(mnorm4, mnorm5) #good to remove 3-way interaction
mnorm6 <- update(mnorm5, .~. -log_shts:habitat)
mnorm7 <- update(mnorm5, .~. -log_shts:shts_0_1)
mnorm8 <- update(mnorm5, .~. -habitat:shts_0_1)
mnorm9 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:shts_0_1)
mnorm10 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:habitat)
mnorm11 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:habitat -log_shts:shts_0_1)
AICtab(mnorm5, mnorm6, mnorm7, mnorm8, mnorm9, mnorm10, mnorm11) #good to remove all interactions except size by density

Anova(mnorm10)
fixef(mnorm10)
```

**No longer significant**

marginally significant size x conspecific density interaction (p = 0.046). No significant main effect of density.

In the GLMM, only plant ID number was retained as a random effect. Plant size (log(shoot number)) was a significant predictor of plant size in the next year ($\chi^2$ = 1422.4, df = 1, P \< 0.001). Conspecific density within 3m of the focal plant did not have a significant effect of plant size in the next year ($\chi^2$ = 1.26, df = 1, P = 0.2623).

```{r glmer-plot, fig.cap="this"}

ha_sub <- ha %>% filter(!is.na(shts_0_1), !is.na(log_shts))
newdata <-
  tibble(log_shts = mean(ha_sub$log_shts),
         habitat = "1-ha", 
         ha_id_number = "new",
         shts_0_1 = seq(min(ha_sub$shts_0_1, na.rm = TRUE),
                          max(ha_sub$shts_0_1, na.rm = TRUE),
                          length.out = 50))

augment(mnorm11, newdata = newdata, re.form = NA) %>% 
  ggplot(aes(x = shts_0_1, y = exp(.fitted))) +
  geom_jitter(data = ha_sub, aes(x = shts_0_1, y = shts_next), alpha = 0.2) +
  geom_line(color = "blue") +
  # geom_ribbon() + #don't have CI yet.  Need to do something like bootMer to get them with bootstrapping.
  labs(x = "shoot density within 3m (shoots/m^2)", y = "shoots (t+1)") +
  theme_bw()
```

# Discussion

We found no meaningful statistically significant effect of conspecific density on the growth or survival of *H. accuminata.* In the GAM approach, the statistically significant effect we saw was driven by a few (\< 20) plants with with a high density of conspecific shoots. The GAM model also did not take into account variation in growth among individual plants, which was shown to be important in the GLMER model. In the GLMER, the effects of density are not significant. Because of this, we feel it is safe to exclude density dependence from further demographic models.

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::