### 2021-02-12 analysis/paper/EDA_figure_1.Rmd
moved figure 1 and data wrangling out to separate files.

---
title: "Exploratory Data Visualizations (Fig. 1)"
author: "Eric R. Scott"
date: "2021-02-12"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(patchwork)
library(HeliconiaDemography) #load functions in current package.
library(conflicted)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

What is the goal of this notebook?

# Load Data

```{r}
demog_full <- read_rds(here("analysis", "data", "derived_data", "ha_spei_combined.rds"))
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
```

## Tidy

There are plants with 0 shoots and 0 ht.  I will remove those, as having no above ground biomass is really not a size, but a different category.
There's also a couple of plants that are probably data entry errors in height.

```{r}
demog_full <-
  demog_full %>% 
  filter(ht < 200 | is.na(ht), ht_prev < 200 | is.na(ht_prev)) %>%
  filter(
    # having no aboveground biomass is possible, but categorically different than just being small. Exclude plants with 0 shoots or 0 height
    shts_prev > 0,
    ht_prev > 0,
    #for survival data, must include plants with NA for shts and height (i.e. dead plants)
    shts > 0 | is.na(shts), 
    ht > 0 | is.na(ht)
  ) %>%
  mutate(size = shts*ht, size_prev = shts_prev * ht_prev, log_size = log(size), log_size_prev = log(size_prev))
```

Only do post-seedlings

```{r}
demog_post <- 
  demog_full %>% 
  filter(code_notes != "sdlg (1)" | is.na(code_notes))
```

# Inspect raw data

```{r}
demog_plotdf <-  
  demog_post %>% 
  filter(habitat !="10-ha") %>% 
  select(ranch, bdffp_reserve_no, plot, habitat, year, ht, shts, shts_prev, size, surv, flwr) %>% 
  mutate(date = paste(year, "-02-01") %>% ymd(),
         yearmonth = yearmonth(date))
```
## SPEI

```{r}
dates <- range(demog_post$year) %>% paste0(., "-02-01") %>% ymd()
census_dates <- min(demog_post$year)[1]:max(demog_post$year) %>% paste(., "-02-01") %>% ymd()

spei <-
  xa %>%
  select(latlon, yearmonth, spei) %>%
  filter(!is.na(spei)) %>% 
  ggplot(aes(x = yearmonth, y = spei)) + 
  geom_line(aes(group = latlon), alpha = 0.2) +
  stat_summary(geom = "line", fun = "mean") +
  scale_y_continuous("SPEI", expand = expansion(mult = c(0, 0.05), add = 0)) +
  scale_x_yearmonth("Date", limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spei <- annotate_spei(spei)
spei
```
## Survival

```{r}
demog_post %>% ggplot(aes(x = surv)) + geom_bar() + facet_wrap(~habitat)
```

```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(surv = sum(surv, na.rm = TRUE) / n())
```

Survival nearly identical for 1 ha and CF.

### Timeseries

```{r}
survival <-
  demog_plotdf %>%  
  filter(shts_prev >= 4) %>%
  ggplot(aes(x = yearmonth, y = surv, color = habitat)) +
  stat_summary(geom = "line", fun = "mean", fun.args = list(na.rm = TRUE)) +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(survived)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
survival
# (survival / spei) 
```
Looks like survival went down year after 2003 drought and during 2007 drought.

## Size / growth

```{r}
demog_post %>% ggplot(aes(x = habitat, y = log_size)) + geom_boxplot()
```
```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(size = mean(size, na.rm = TRUE),
            shts = mean(shts, na.rm = TRUE),
            ht = mean(ht, na.rm = TRUE))
```

### Timeseries

```{r}
size <-
  demog_plotdf %>% 
  ggplot(aes(x = yearmonth, y = size, color = habitat)) +
  stat_summary(geom = "line", fun = "mean", fun.args = list(na.rm = TRUE)) +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("Size") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
size
# size/survival/spei
```
Average plant size shrank in 2003 drought, but not 2007 drought

## Flowering

```{r}
demog_post %>% ggplot(aes(x = flwr)) + geom_bar() + facet_wrap(~habitat)
```
### Timeseries

```{r}
flowering <-
  demog_plotdf %>% 
  filter(shts >= 4) %>% 
  ggplot(aes(x = yearmonth, y = flwr, color = habitat)) + 
  stat_summary(geom = "line", fun = "mean") +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(flowering)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
flowering
```
```{r}
demog_plotdf %>% filter(year == 2003) %>% group_by(habitat) %>% summarize(flwr = mean(flwr, na.rm = TRUE))
```

```{r}
size/survival/flowering/spei +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(plot.margin = margin(2,1,1,1)) &
  guides(col = guide_legend(title = "Habitat"))
ggsave(here("analysis", "figures", "eda.png"), height = 7, width = 10)
```