### 2021-02-26 analysis/paper/EDA_figure_1.Rmd
used coord_cartesian to set limits instead of scale.  Dodged point was missing.

---
title: "Exploratory Data Visualizations (Fig. 1)"
author: "Eric R. Scott"
date: "2021-02-12"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, paged.print = FALSE)
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(patchwork)
library(HeliconiaDemography) #load functions in current package.
library(conflicted)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

What is the goal of this notebook?

# Load Data

```{r}
demog <- read_rds(here("analysis", "data", "derived_data", "ha_spei_combined.rds"))
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
```

## Tidy

There are plants with 0 shoots and 0 ht.
I will remove those, as having no above ground biomass is really not a size, but a different category.
There's also a couple of plants that are probably data entry errors in height.

```{r}
demog_full <-
  demog %>% 
  filter(ht < 200 | is.na(ht), ht_prev < 200 | is.na(ht_prev)) %>%
  filter(
    # having no aboveground biomass is possible, but categorically different than just being small. Exclude plants with 0 shoots or 0 height
    shts_prev > 0 | is.na(shts_prev),
    ht_prev > 0 | is.na(shts_prev),
    #for survival data, must include plants with NA for shts and height (i.e. dead plants)
    shts > 0 | is.na(shts), 
    ht > 0 | is.na(ht)
  ) %>%
  mutate(size = shts*ht, size_prev = shts_prev * ht_prev, log_size = log(size), log_size_prev = log(size_prev))
```

Only do post-seedlings

```{r}
demog_post <- 
  demog_full %>% 
  filter(code_notes != "sdlg (1)" | is.na(code_notes))
```

# Inspect raw data

```{r}
demog_plotdf <-  
  demog_post %>% 
  filter(habitat !="10-ha") %>% 
  select(ranch, bdffp_reserve_no, plot, habitat, year, ht, shts, shts_prev, size, surv, flwr) %>% 
  mutate(date = paste(year, "-02-01") %>% ymd(),
         yearmonth = yearmonth(date))
```

## SPEI

```{r}
dates <- 
  range(demog_post$year) %>%
  paste0(., "-02-01") %>% 
  ymd()
dates[1] <- dates[1] - months(36)

census_dates <- seq(dates[1], dates[2], by = "year")

spei <-
  xa %>%
  select(latlon, yearmonth, spei) %>%
  filter(!is.na(spei)) %>% 
  ggplot(aes(x = yearmonth, y = spei)) + 
  geom_line(aes(group = latlon), alpha = 0.2) +
  stat_summary(geom = "line", fun = "mean") +
  scale_y_continuous("SPEI", expand = expansion(mult = c(0, 0.05), add = 0)) +
  scale_x_yearmonth("Date", limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spei <- annotate_spei(spei)
spei
```

## Survival

```{r}
demog_post %>% ggplot(aes(x = surv)) + geom_bar() + facet_wrap(~habitat)
```

```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(surv = sum(surv, na.rm = TRUE) / n())
```

Survival nearly identical for 1 ha and CF.

When is survival lowest?

```{r}
demog_post %>% 
  group_by(year, habitat) %>% 
  summarize(surv = sum(surv, na.rm = TRUE) / n()) %>% 
  pivot_wider(names_from = habitat, values_from = surv)
```

### Timeseries

```{r}
survival <-
  demog_plotdf %>%  
  filter(shts_prev >= 4) %>%
  ggplot(aes(x = yearmonth, y = surv, color = habitat)) +
  stat_summary(geom = "line", fun = "mean", fun.args = list(na.rm = TRUE)) +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(survived)") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "none"
  )
survival
# (survival / spei) 
```

Looks like survival went down year after 2003 drought and during 2007 drought.

### Survivorship curve

Emilio suggested an alternative to the timeseries plot---a survivorship curve for the 1998 cohort of plants.

```{r paged.print=FALSE}
cohort <- demog %>% filter(year == 1998) %>% pull(ha_id_number) %>% unique()

surv_curve_df <-
  demog %>% 
  filter(ha_id_number %in% cohort) %>% 
  filter(surv == 1, habitat != "10-ha") %>% 
  group_by(year, habitat) %>% 
  summarize(n = n()) %>%
  group_by(habitat) %>% 
  mutate(p_surv = n/first(n),
         date = ymd(paste0(year, "-02-01")),
         yearmonth = yearmonth(date))

surv_curve <-
  ggplot(surv_curve_df, aes(x = yearmonth, y = p_surv, color = habitat)) +
  geom_step() +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.04)) +
  scale_y_continuous("P(survived)") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
surv_curve
```

## Size / growth

```{r}
demog_post %>% ggplot(aes(x = habitat, y = log_size)) + geom_boxplot()
```

```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(across(
    c(size, shts, ht),
    list(mean = mean, sd = sd),
    na.rm = TRUE,
    .names = "{.col}_{.fn}"
  ))

```

### Timeseries

```{r}
size <-
  demog_plotdf %>% 
  ggplot(aes(x = yearmonth, y = log(size), color = habitat)) +
    stat_summary(
      geom = "line",
      fun = "mean",
      fun.args = list(na.rm = TRUE),
      position = position_dodge(width = 100),
      aes(group = habitat)
    )+
    stat_summary(
      geom = "pointrange",
      fatten = 1,
      fun.data = "mean_sdl",
      fun.args = list(mult = 1),
      position = position_dodge(width = 100)
    ) +
  scale_x_yearmonth(
    # limits = dates,
    breaks = census_dates,
    date_minor_breaks = "1 month",
    expand = expansion(mult = 0.04)
  ) +
  scale_y_continuous("log(size)") +
  coord_cartesian(xlim = dates) + #limits here so points don't get removed
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "none"
  ) +
  NULL
# size/survival/spei
```

Average plant size shrank in 2003 drought, but not 2007 drought

### Ridge plot

```{r}
library(ggridges)
ggplot(demog_plotdf) +
  geom_density_ridges(aes(x = log(size), y = as.factor(year), fill = habitat, color = habitat), alpha = 0.5)
ggplot(demog_plotdf) +
  geom_violin(aes(y = log(size), x = as.factor(year), fill = habitat), alpha = 0.5, draw_quantiles = c(0.5))

```

## Flowering

```{r}
demog_post %>% ggplot(aes(x = flwr)) + geom_bar() + facet_wrap(~habitat)
```

How many infloresences given flowering?

```{r}
demog %>% 
  filter(flwr == 1) %>%
  summarize(mean_infl = mean(infl_num),
            median_infl = median(infl_num),
            min = min(infl_num),
            max = max(infl_num))
```

### Timeseries

```{r}
flowering_df <-
  demog_plotdf %>% 
  filter(shts >= 4) %>% 
  group_by(yearmonth, habitat) %>% 
  summarize(n = n(), flwr = sum(flwr == 1, na.rm = TRUE)) %>% 
  mutate(prop_flwr = flwr/n)

flowering <-
  ggplot(flowering_df, aes(x = yearmonth, y = prop_flwr, color = habitat)) + 
  geom_line() +
    scale_x_yearmonth(
      limits = dates,
      breaks = census_dates,
      date_minor_breaks = "1 month",
      expand = expansion(mult = 0.04)
    ) +
    scale_y_continuous("P(flowering | shts ≥ 4)") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
flowering
```

```{r}
demog_plotdf %>% filter(year == 2003) %>% group_by(habitat) %>% summarize(flwr = mean(flwr, na.rm = TRUE))
```

# Combine

```{r}
size/
  # survival/
  surv_curve/
  flowering/
  spei +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(plot.margin = margin(2,1,1,1)) &
  guides(col = guide_legend(title = "Habitat"))
ggsave(here("analysis", "figures", "eda.png"), height = 7, width = 10)
```