### 2021-03-01 analysis/paper/paper.Rmd
Changes in Methods and Introduction UP TO the "interaction" section.
saving here to make it easier to revert if unhappy with new intro focusing on hab diff with no frag x xlimate interaction).

---
title: "Delayed effects of [responses to?] climate lead to demographic divergence in a fragmented Amazonian landscape"
author:
  - Eric R. Scott:
      email: scottericr@gmail.com
      institute: [UF_WEC]
      correspondence: true
  - Maria Uriarte:
      email: mu2126@columbia.edu
      institute: [Columbia]
      correspondence: false
  - Emilio M. Bruna:
      email: embruna@ufl.edu
      intitute: [UF_WEC,UF_LAS,BDFFP]
      correspondence: false

institute:
  - UF_WEC: Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, Florida 32611-0430
  - UF_LAS: Center for Latin American Studies, University of Florida, Gainesville, Florida 32611-5530  
  - Columbia: Department of Ecology, Evolution and Environmental Biology, Columbia University 1200 Amsterdam Avenue, New York, New York 10027 USA
  - BDFFP: Biological Dynamics of Forest Fragments Project, INPA-PDBFF, CP 478, Manaus, Amazonas 69011-970 Brazil
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      number_sections: no
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
linkcolor: blue
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  forest fragmentation; vital rates; *Heliconia acuminata*; drought; ENSO; edge effects
highlights: |
  These are the highlights
editor_options:
  markdown:
    wrap: sentence
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = "../figures/",
	message = FALSE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>",
	dpi = 300
)
library(knitr)
library(here)
```

<!-- The actual document text starts here: -->

# Introduction

<!-- Underemphasize the interaction between fragmentation and drought bc can't do the interaction in test. Instead it's "compare drought effects in fragmented and cf populations."  -->
<!-- In response to this I deleted #1; not looking at interaciton of multiple effects and not studying population dyanmics per se in this one. Suggest alluding to this in Discussion instead. -->

 <!-- [1]. Although the local extinction of species from fragments is common [1], the demographic mechanisms underlying these extinctions are rarely known [2]. However, it is often assumed – especially in the tropics – that these extinctions are driven by the dramatically altered abiotic conditions in fragments [3] reducing rates of reproduction or survivorship [4-6]. This is why understanding the way in which landscape structure and environmental conditions interact to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [7,8].  -->

Deforestation driven by the expansion of agriculture and other human activities is a primary driver of biodiversity loss worldwide [@haddadHabitatFragmentationIts2015]. Deforestation also results in landscapes where the remaining forest can be highly fragmented, with remnants of different sizes embedded in a matrix of often highly-contrasting habitat. Although the local extinction of species from individual fragments is common *(CITATIONS_1)*, the demographic mechanisms underlying these extinctions are poorly understood *(CITATIONS_2)*. However, it is often assumed that that the dramatically altered abiotic conditions often observed in fragments can depress rates of reproduction or survivorship *(CITATIONS_3)*, ultimately leading to local or regional population declines. This is thought to be particularly true in species-rich tropical forests, which continue to be cleared and fragmented at unprecedented rates. For example, over 24 million ha of the Brazilian Amazon have been cleared in the last two decades *(CITATION_4)*, resulting in extensive fragmentation and creation of over 70,000 km of new forest edges annually [@broadbentForestFragmentationEdge2008].

<!-- CITATIONS_1: Haddad, others? -->
<!-- CITATIONS_2: Bruna et al J Veg Sci 2009, others? -->
<!-- CITATIONS_3: Bruna 1999 Nature, Laurance et al 1998 Ecology, Zartman 2015 Biotropica.  -->
<!-- CITATION_4: WRI Report? -->
<!-- CITATION_5: Brodie J et al 2012. TREE, Scheffers et al 2016. Science -->
<!-- CITATION_6: Zeng et al 2008 Environmental Research Letters, Mora et al 2013 Nature -->

Populations in fragmented tropical landscapes are also at risk from climate change *(CITATIONS_5)*. These effects may be especially pronounced in Amazonia, parts of which have been experiencing a drying trend since the mid-1970s and where climate models for the 21st century predict substantial decreases in dry season precipitation [@malhi2008] as well as increased drought frequency and severity [@caiIncreasingFrequencyExtreme2014] *(add CITATIONS_6)*.
Severe droughts in the Amazon have been shown to reduce aboveground biomass through increased tree mortality and dieback [@williamsonAmazonianTreeMortality2000;[@phillipsDroughtSensitivityAmazon2009; @marengoDroughtAmazonia20052008; @lewis2010AmazonDrought2011].

4.  It has been hypothesized that drought and fragmentation could act in synergy to further reduce population viability in fragmented landscapes, particularly in the tropics [@lauranceEffectsStrongDrought2001; @didhamEdgeStructureDetermines1999].
    Fragmentation drives several changes in microclimate near forest edges, including increased air temperature, decreased relative humidity, reduced soil moisture, and increased penetration of solar radiation in the forest understory [@ewersFragmentationImpairsMicroclimate2013; @didhamEdgeStructureDetermines1999].
    This decrease in microclimatic buffering could exacerbate the effects of drought in fragments compared to intact forest.
    This may be why the drought-induced mortality of trees is higher near forest edges than in forest interiors [@schwartz2019; @lauranceEffectsStrongDrought2001].
    Alternatively, plant populations persisting in fragments may be more resilient to drought if individuals susceptible to desiccation died quickly after fragment isolation [@betts2019].

5.  The effects of drought and fragmentation could differ among demographic vital rates such as growth, survival, and reproduction.
    In addition to variation among vital rates, responses to drought and fragmentation could depend on life history stage or plant size.
    Large established adults could respond differently than seedlings.
    This is unknown because there are very few systems where we have data on entire life history in a climate change or fragmentation context.
    <!--# I'm not actually testing a size x climate or size x fragmentation interaction explicitly.  Maybe I should not emphasize the variation among size classes in the introduction -->

    -   Fragmentation affects plant size (in *Heliconia* [@brunaDemographicEffectsHabitat2005] and trees [@schwartz2019]). Smaller/larger plants might be more susceptible to drought. So differential effects of drought in fragments and continuous forest could be due to an interaction between drought and plant size.

To date, however, this hypothesis remains untested for three primary reasons.
First, most studies elucidating how tropical plants respond to fragmentation have studied either juvenile (i.e., seed, seedling) or larger, reproductive plants.
<!--# example of juvenile but not reproductive study with drought -->.
<!--# example of reproductive but not juvenile study with drought -->.
Additionally, while most studies of tropical plant demography focus on trees, little is known about the understory plants that comprise up to \_\_\_\_% of biomass in tropical forests and are ecologically important <!--# citations -->.

Second, there is a growing literature on how tropical plants respond to droughts, but few studies have compared the responses of plants in continuous forest with those in forest fragments.
Laurance et al. [-@lauranceEffectsStrongDrought2001] found that the increase in tree (\>10 cm DBH) mortality during a drought were exacerbated in fragments, but the decline in mortality after the drought was also stronger in fragments.
However, there are no similar studies on tropical understorey plants.

Finally, the long-term data needed to test population-level hypotheses about climate change-fragmentation synergies are scant, especially for tropical systems [@salguero-gomezCOMPADREPlantMatrix2015; @croneHowPlantEcologists2011].
Without multi-decadal sampling one cannot capture enough droughts to quantify their ecological impact, nor the variation in vital rates needed to parameterize demographic models of long-lived species in changing climates [@morrisQuantitativeConservationBiology2002; @teller2016].
Long-term data are also needed because while some demographic effects of fragmentation or drought on can be detected immediately, others may take years to manifest.
Lagged effects of climate on plant vital rates may, in fact, be the rule rather than the exception [@eversLaggedDormantSeason].

We investigated the effects of drought on the growth, survival, reproduction, and recruitment <!--# I haven't included recruitment.  Should I? --> of a understory herb (*Heliconia acuminata*) in an experimentally fragmented landscape in the Central Amazon.
Specifically, we asked 1) Does drought increase or decrease the growth, survival, and fertility rates of plant populations in continuous forest?
2) What lag times are most important in considering the effects of drought on vital rates?
3) Are the effects of drought on the vital rates of populations in fragments similar in direction and magnitude to those in continuous forest?

# Methods

```{r}
cap <- "This is just a placeholder figure, but a map or satelite image of BDFFP will go here."
```

```{r map, fig.cap=cap}
include_graphics(here("analysis", "figures", "map.png"))
```

## Study site

The Biological Dynamics of Forest Fragments Project (BDFFP) is located \~70 km north of Manaus, Brazil (2º30' S, 60ºW). In addition to large areas of continuous forest, the BDFFP has forest fragment reserves isolated from 1980--1984 by felling the trees surrounding the area chosen for isolation and, in most cases, burning the downed trees once they dried (Figure \@ref(fig:map). In subsequent decades the vegetation regenerating around fragments has been periodically cleared to ensure fragment isolation [@bierregaardLessonsAmazoniaEcology2001].

The BDFFP reserves are located in nonflooded (i.e., *terra firme*) tropical lowland forest with a 30--37m tall canopy [@rankin-de-meronaPreliminaryResultsLargescale1992] and an understory dominated by stemless palms [@scariotForestFragmentationEffects1999]. The soils in the reserves are nutrient-poor xanthic ferrosols; their water retention capacity is poor despite having a high clay content. Mean annual temperature in the region is 26º C (range=19--39º C), and annual rainfall at the BDFFP reserves ranges from 1900-2300 mm. There is a pronounced dry season from June to October **(Appendix?)**.

## Focal species

*Heliconia acuminata* (LC Rich.) (Heliconiaceae) is a perennial monocot distributed throughout Central Amazonia [@kressDiversityDistributionHeliconia1990] and is the most abundant understory herb at the BDFFP [@ribeiroInfluencePostclearingTreatment2010]. While many *Heliconia* species grow in large patches along forest edges or in treefall gaps, others---such as *H. acuminata*---are found at lower densities in the darker and cooler forest understory. These species produce fewer infloresences and are pollinated by traplining rather than territorial hummingbirds [@berryHeliconiaIdentificationGuide1991a]. While some *Heliconia* species are at least partially self-compatible *[Kress Evolution 1983]*, pollen transfer experiments indicate self-fertilization in *Heliconia acuminata* is extremely rare *[darrigo unpubl. data]*.

In our sites *H. acuminata* is pollinated by *Phaeothornis superciliosus* and *P. bourcieri* [@bruna2004]. Plants begin flowering at the start of the rainy season; reproductive plants have $\bar{x}$ = **1.1** flowering shoots (range = 1-7), each of which has an inflorescence with 20--25 flowers [@brunaHabitatFragmentationDemographic2002]. Fruits mature April-May, have 1--3 seeds per fruit ($\bar{x}$=2), and are eaten by a thrush and several species of manakin [@uriarteDisentanglingDriversReduced2011]. Dispersed seeds germinate \~6 months after dispersal at the onset of the subsequent rainy season, with rates of germination and seedling establishment higher in continuous forest than forest fragments [@brunaHabitatFragmentationDemographic2002; @brunaSeedGerminationRainforest1999].

## Demographic data collection

In 1997--1998 we established a series of 5000 $m^{2}$ plots ($50 \times 100$m) in the BDFFP reserves in which we marked and measured all *Heliconia acuminata*. The plots are located in continuous forest (n=6), 10-ha fragments (n=3), and 1-ha fragments (n=4), with distance between plots ranging from 500 m--41 km. Plots in 1-ha fragments were on one randomly selected half of the fragment, plots in 10-ha fragments were in the fragment center, and plots in continuous forest were placed in locations 500-4000 m from the borders of secondary and mature forest (Fig or Table <!--# insert fig ref -->). This study uses data from 1-ha fragments and Continuous Forest sites.

Each plot is subdivided into 50 quadrats ($10 \times 10$m) to simplify the annual surveys, during which we recorded the number of vegetative shoots each plant had, the height of each plant to the tallest leaf, and the identity of any flowering plants. We also surveyed plots regularly during the rainy season to identify any that flowered after the survey. Height and shoot number are correlated with leaf area, the probability of flowering, and rates of survivorship [@brunaHabitatFragmentationDemographic2002; @brunaEffectsForestFragmentation2002]. Plants that were not found for three consecutive surveys were considered dead. There is a seven-fold range in *H. acuminata* density in the forest plots, with density in several of the plots comparable to that in low-density fragments [@brunaHabitatFragmentationDemographic2002].

## Weather data

Data on precipitation and potential evapotranspiration in our sites were obtained from a published gridded dataset (0.25º x 0.25º resolution) built using data from 3,625 ground-based weather stations across Brazil [@xavierDailyGriddedMeteorological2016].
We used these data to calculate the standardized precipitation evapotranspiration index (SPEI) [@vicente-serranoMultiscalarDroughtIndex2010].
SPEI is a proxy for meteorological drought that integrates precipitation and evapotranspiration anomalies over a specified time scale.
Positive SPEI values for a given month indicate conditions wetter than the historical average for that month, while negative values of SPEI indicate droughts with intensity categorized as mild (0 to -1), moderate (-1 to -1.5), severe (-1.5 to -2), or extreme (\< -2) [@mckeeRelationshipDroughtFrequency1993].
SPEI can be calculated to represent different temporal scales of drought; we used 3-month SPEI because -- given its shallow roots and rhizome -- *H. acuminata* relies primarily on soil moisture rather than deeper water sources that can persist for longer timescales.
SPEI calculations were done using the `SPEI` package in `r version$version.string` [@rcoreteamLanguageEnvironmentStatistical2020; @begueriaSPEICalculationStandardised2017].
The timing of drought events based on these SPEI calculations is consistent with that resulting from SPEI calculated with other data sources (e.g., nearby weather stations, the global SPEI database [@begueriaSbegueriaSPEIbaseVersion2017], TRMM [@trmm3b43]), though the magnitude of drought can sometimes differ (**Appendix \_\_**).

## Statistical Modeling of _______________ 
<!-- I think this is well described but could use some biologizing, eg,"This function allows the effect of a predictor on a response variable (e.g., SPEI on the growth rate of plants) to vary non-linearly through lag time and across predictor intensity (e.g, SPEI values from -2 to 2).  -->
We assessed the effects of drought history on plant vital rates using distributed lag non-linear models (DLNMs) [@gasparriniPenalizedFrameworkDistributed2017].
DLNMs model potentially delayed effects of predictor variables by fitting a bi-dimensional predictor-lag-response association spline, referred to as a crossbasis function.
This function allows the effect of a predictor to vary non-linearly through lag time and across predictor intensity.
Crossbasis functions were created using the *dlnm* package with possible lags from 0--36 months [@gasparriniDistributedLagLinear2011; @rcoreteamLanguageEnvironmentStatistical2020].
We chose 36 months as a maximum lag because prior *H. acuminata* individuals transplanted to forest fragments typically recovered from transplant shock within 36 months [@brunaExperimentalAssessmentHeliconia2002].
The crossbasis function was fit to the data in the context of a generalized additive model (GAM) using the *mgcv* package with restricted maximum likelihood [@woodGeneralizedAdditiveModels2017].
The general form of the vital rate models was as follows:

$$
\begin{gather}
    g[ \textrm{E}(y_{i})] = \alpha_j + s_1(x_i) + s_2(d_{i,t}, ..., d_{i,t-24}) \\
    \alpha_j \sim N(\mu_{\alpha_j}, \sigma^2_{\alpha_j}), \textrm{for plot }j 
\end{gather}
$$

Where $s_1(x_i)$ is a smooth function of plant size [ ln(height $\times$ shoot number) ], fit using a penalized cubic regression spline, $s_2(\cdot)$ is the crossbasis function in which $d_{i,t}$ is the SPEI value during the census month of an observation (February) and $d_{i,t-l}$ is the SPEI $l$ months prior (see Gasparrini et al. 2017 for details).
The crossbasis function, $s_2(\cdot)$ can also be written:

$$
s_2(d_{i,t}, ..., d_{i,t-24}) = \sum_{l = l_0}^L f \cdot w(d_{i,t-l},l) 
$$

Where the crossbasis function, $f \cdot w (d,l)$, is composed of two marginal basis functions: the standard predictor-response function $f(d)$, and the additional lag-response function $w(l)$.
These marginal functions are combined as a tensor product smooth such that the shape of one marginal function varies smoothly along the other dimension (see Wood [-@woodsimonn.ChapterSmoothers] and Gasparrini et al. [-@gasparriniPenalizedFrameworkDistributed2017] for more detail).
Penalized cubic regression splines were used for both marginal bases of the crossbasis function, with 35 knots for the lag dimension and 3 knots for the drought response dimension to restrict the fitted response to drought to bimodal at most complex.
A random effect of plot ID on the intercept is represented by $a_j$.
For plant growth the response was plant size (natural log of height $\times$ shoot number) in year t+1 modeled with a scaled t family error distribution.
A scaled t family was chosen because residuals were leptokurtic with a Gaussian error structure.
The data on number of infloresences was converted to binary (1 for ≥1 inflorescence, 0 for no inflorescences).
When survival and flowering were the response variables we used a binomial family error distribution with a logit link function.

In the process of fitting the models, the penalty on the crossbasis smooth (and other smoothed terms) is optimized such that more linear shapes are favored unless the data supports non-linearity [@woodGeneralizedAdditiveModels2017].
We applied an additional penalty to shrink linear portions toward zero with the `select=TRUE` option to the `gam()` function, and inferred statistical significance of model terms with p-values from the `summary.gam()` function as recommended in Marra and Wood [-@marraPracticalVariableSelection2011].

Because the `dlnm` package does not currently allow modeling of factor by smooth interactions, we fit separate models for plants in fragments and in continuous forest to allow the shape of the crossbasis function to differ between habitats.

# Results

```{r}
cap <- "Timeseries of demographic parameters and drought ocurrence. Mean plant size ± standard deviation (a) shows that plants were smaller in fragments (red) than continuous forest (blue). (b) shows the survival curve for the 1998 cohort of plants in both fragments and continuous forest. c) shows the proportion of mature plants (with 4 or more shoots) flowering each census. In d), monthly 3-month SPEI is plotted. Light grey lines represent values from different grid cells encompassing BDFFP and the dark line is the site mean.  Yellow, orange, dark orange, and red stripes show mild, moderate, severe, and extreme drought, respectively."
```

```{r eda, fig.cap=cap}
knitr::include_graphics(here("analysis", "figures", "eda.png"))

```

The meteorological drought occurrences indicated by SPEI mostly agree with drought events in the Amazon reported in the literature.
For example, the drought associated with the 1997 El Niño Southern Oscillation (ENSO) event was one of the most severe on record for the Amazon [@williamsonAmazonianTreeMortality2000]; correspondingly, 1997 has the lowest SPEI values in our timeseries (Figure \@ref(fig:eda)d).\
The 2005 dry season (June--October) was also exceptionally dry, although this drought mostly affected the south western Amazon [@zengCausesImpacts20052008; @marengoDroughtAmazonia20052008]. <
!--# I know I'm missing some literature here.  Add more refs for more droughts. Any papers mentioning 2003 wet season drought? Also, Feb 2007?-->

Mean plant size dropped dramatically in 2003 in both habitat types, corresponding with the most severe drought in the timeseries (Figure \@ref(fig:eda)d).
Two noticeable peaks in the proportion of flowering mature plants occur in the most severe drought years (2003 and 2007) (Figure \@ref(fig:eda)c).

## Survival

```{r}
cap <- "Smooth effect of plant size in the previous census on survival modeled by a cubic regression spline.  95% confidence intervals are shown and include uncertainty in the model intercept and uncertainty due to smoothness selection. The smooths for 1-ha fragments and continuous forest are fit in separate models. The model intercepts were added to the fitted values of the smooth and values are back-transformed to the response scale"
```

```{r surv-size, fig.cap=cap}
include_graphics(here("analysis", "figures", "surv_both_size.png"))
```

Overall survival rate across the entire dataset is 0.95 in both CF and 1-ha.
The proportion of surviving plants across both habitat types was lowest in 2004 (0.92) which coincided with a moderate drought in Jan 2004 / Feb 2005 and a severe drought in Jan--Mar of 2004.
The lowest survival for 1-ha fragment plots (0.93) was in the 2006 census corresponding to unusually wet weather (SPEI \>0.5), but with a moderate drought the preceding October.

Log transformed plant size had a significant effect on survival in both continuous forest and 1-ha fragments (both p \< 0.001).
Survival probability of large plants approached 1 in both habitat types (Figure \@ref(fig:surv-size)).
Survival of the smallest plants was higher in 1-ha fragments, but the 95% confidence intervals for 1-ha fragments and continuous forest overlapped for all sizes.

```{r}
cap <- "Smooth effect of lagged SPEI on survival in a) continuous forest and b) 1-ha fragments. Surface is modeled as a crossbasis function with cubic regression splines for each marginal basis. Model intercepts were added to fitted values of the crossbasis function and back-transformed to the response scale. Contour lines correspond to a change of 0.002. The bar on the bottom of each panel indicates wet seasons (black, November–May) and dry seasons (white, June–October)."

```

```{r surv-spei, fig.cap=cap}
knitr::include_graphics(here("analysis", "figures", "surv_both_spei.png"))
```

Drought history had a significant effect on survival in both habitats (P \< 0.001), but the shape of the crossbasis surface differed.
For 1-ha fragments, recent SPEI values had the strongest effect on survival with the highest survival near SPEI of 0 and decreasing survival in both drier and wetter conditions (Figure \@ref(fig:surv-spei)a).
In continuous forest, recent SPEI had a weaker effect with only SPEI values less than -1 producing slightly reduced survival probabilities (Figure \@ref(fig:surv-spei)b).
However, SPEI at a lag time of 15--20 months, corresponding to June--October of the previous year (i.e. not the most recent dry season, but the one before that), also effected survival probabilities.
Lower SPEI (drought) 15--20 months prior to census was associated with reduced survival and higher SPEI at these lag times was associated with higher survival.
Plants in both continuous forest and 1-ha fragments showed an increase in survival probability with very high (wet) SPEI values at a lag time of 36 months.
However, it should be noted that only the first year of census data (1999) had an SPEI value greater than 2 within 36 months in the past.

```{r}
cap <- "The difference between predicted effects of SPEI history on survival in continuous forest and 1-ha fragments."
```

```{r surv-diff, fig.cap=cap}
include_graphics(here("analysis", "figures", "surv_diff_spei.png"))
```

Subtracting the estimated contribution of SPEI history to survival in fragments from that in continuous forest shows that in average conditions (SPEI = 0), there is little difference in survival probability between continuous forest and fragments (Figure \@ref(fig:surv-diff)).
However, under extreme conditions, survival probability is higher in continuous forest by up to 0.02.

The random effect of plot was significant for continuous forest ( p \< 0.001), but not for 1-ha fragments (p = 0.257).

## Growth

Plants were on average larger in continuous forest compared to 1-ha fragments.
<!--# This is across the whole timeseries, which I'm not sure is actually that important. What summary statistics would be useful? -->In continuous forest, plants had an average of 2.9 ± 1.8 shoots (mean ± standard deviation) and were on average 40.6 ± 26.5 cm tall with an average size (shoots $\times$ height) of 150 ± 175.
In 1-ha fragments, plants had 2.5 ± 1.5 shoots on average and a mean height of 36.2 ± 24.1 cm and a mean size of 112 ± 141.
The size difference between plants in continuous forest and 1-ha fragments was more dramatic early on in the timeseries with mean sizes becoming more similar over time (Figure \@ref(fig:eda)a).

```{r}
cap <- "Smooth effect of plant size in year t on plant size in year t+1 modeled by a cubic regression spline.  95% confidence intervals are shown and include uncertainty in the model intercept as well as uncertainty due to smoothness selection. The dashed red line represents a 1:1 line for reference. The smooths for 1-ha fragments and continuous forest are fit in separate models. Model intercepts were added to the fitted values of the smooths."
```

```{r growth-size, fig.cap=cap}
include_graphics(here("analysis", "figures", "growth_both_size.png"))
```

Log transformed size in year t was a significant predictor of log transformed size in year t+1 (p \< 0.001).
The 95% confidence intervals for the effect of size in year t in fragments and continuous forest overlapped for most sizes, but was slightly higher in continuous forest for mid-sized plants (Figure \@ref(fig:growth-size)).

```{r}
cap <- "Smooth effect of lagged SPEI on plant growth for a) continuous forest and b) 1-ha fragments. Surface is modeled as a crossbasis function with cubic regression splines for each marginal basis. Model intercepts were added to fitted values of the crossbasis function. Contour lines correspond to a change of 0.05. The bar on the bottom of each panel indicates wet seasons (black, November–May) and dry seasons (white, June–October)."
```

```{r growth-spei, fig.cap=cap}
include_graphics(here("analysis", "figures", "growth_both_spei.png"))
```

Drought history also had a significant effect on plant size in both habitats (p \< 0.001).
The fitted crossbasis functions for fragments and continuous forest showed somewhat similar patterns, although the surface for 1-ha fragments was much more complex (edf = 17.9 for 1-ha fragments; edf = 5.9 for continuous forest; see also Figure \@ref(fig:growth-spei)).
Drought at lags 8--11 (March - June, the end of the wet season) increase growth in both habitats, but more strongly in 1-ha fragments.

```{r}
cap <- "The difference between predicted effects of SPEI history on plant size in continuous forest and 1-ha fragments."
```

```{r growth-diff, fig.cap=cap}
include_graphics(here("analysis", "figures", "growth_diff_spei.png"))
```

The difference between the effects of SPEI history in continuous forest and 1-ha fragments shows that growth is equal or slightly higher in continuous forest for average conditions (SPEI = 0) over all lag periods (Figure \@ref(fig:growth-diff).
However, growth in continuous forests is higher by up to log(size) = 0.48 <!--# double-check this number --> when the current wet season is unusually wet (SPEI\>1).

The random effect of plot is significant in both habitat models (p \< 0.001).

## Flowering

Probability of flowering was generally higher in continuous forest compared to 1-ha fragments.
The exceptions to this are in 2006 and 2008 when the proportion of flowering plants was lower in continuous forest (Figure \@ref(fig:eda)c).
The most striking difference between habitat types in flowering proportion was in 2003 when 9.7% of plants flowered in the continuous forest plots and only 2.7% flowered in fragment plots.
<!--# update to be only for large plants -->

```{r}
cap <- "Smooth effect of plant size in the previous census on flowering probability modeled by a cubic regression spline.  95% confidence intervals are shown and include uncertainty in the model intercept as well as uncertainty due to smoothness selection. The smooths for 1-ha fragments and continuous forest are fit in separate models. The model intercepts were added to the fitted values of the smooth and values are back-transformed to the response scale"
```

```{r flwr-size, fig.cap=cap}
include_graphics(here("analysis", "figures", "flwr_both_size.png"))
```

Plant size had a significant effect on flowering probability in both habitats (p \< 0.001).
Flowering probability increased rapidly with log-transformed plant size greater than 5 in both habitats (Figure \@ref(fig:flwr-size)).
Flowering probability of large plants (log(size) \> 5) was greater in 1-ha fragments compared to continuous forest.
However, because mean plant size is larger in continuous forest, overall flowering probability was greater.
<!--# idea: plot vertical lines representing mean or median size in each habitat -->

```{r}
cap <- "Smooth effect of lagged SPEI on flowering probability. Surface is modeled as a crossbasis function with cubic regression splines for each marginal basis. Model intercepts were added to fitted values of the crossbasis function and back-transformed to the response scale. Contour lines correspond to a change of 0.001. The bar on the bottom of each panel indicates wet seasons (black, November–May) and dry seasons (white, June–October)."
```

```{r flwr-spei, fig.cap=cap}
include_graphics(here("analysis", "figures", "flwr_both_spei.png"))
```

Drought history had a significant effect on flowering probability in both habitat types (p \< 0.001).
Flowering probability was overall higher at all values of SPEI in continuous forest compared to 1-ha fragments (Figure \@ref(fig:flwr-spei)).
The effects of SPEI on flowering probability were also stronger and more complex in continuous forest compared to fragments (1-ha edf = 8.5, continuous forest edf = 10.5).
In continuous forest, flowering probability was increased by recent drought (at lag = 0 with SPEI \< -1) and drought at lags 15--20 (Jun-Nov, the dry season in the previous year).
Drought at lags 7--13 reduced flowering probability.
The shape of the crossbasis function was similar in 1-ha fragments, but significantly muted (Figure \@ref(fig:flwr-spei)a).
That is, the variation in fitted values along the SPEI dimension was much lower at all values of the lag dimension in fragments.

```{r}
cap <- "The difference between predicted effects of SPEI history on flowering probability in continuous forest and 1-ha fragments."
```

```{r flwr-diff, fig.cap=cap}
include_graphics(here("analysis", "figures", "flwr_diff_spei.png"))
```

Plotting the difference between the crossbasis smooths for continuous forest and 1-ha fragments revealed that flowering probability is higher in continuous forest compared to 1-ha fragments and is increased by recent drought and drought during the dry season 15--20 months prior to the census (Figure \@ref(fig:flwr-diff)).
The least difference between habitat types occurs when there is strong drought at about 8--12 months lag or very wet conditions at 15--20 months lag.

The random effect of plot also had a significant effect in both habitats (p \< 0.001).

# Discussion

Understanding how landscape structure and environmental conditions act to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [7,8]. 
<!-- Didham RK, Kapos V, Ewers RM  2012. Rethinking the conceptual foundations of habitat fragmentation research. Oikos 121:161-170. -->
<!-- Driscoll DA, Banks SC, Barton PS, Lindenmayer DB, Smith AL  2013. Conceptual domain of the matrix in fragmented landscapes. Trends in Ecology & Evolution 28:605-613. -->

1. To set up the next paper: Organisms simultaneously experience multiple forms of disturbance. While work investigating the demographic effects of individual factors is common, studies investigating how multiple factors interact to influence population dynamics remains poorly understood. They could act in concert to exacerbate or dampen population dynamics.

Caveats:

<!--# need to edit.  Moved from methods. -->

One limitation of the DLNM approach is that it does not currently allow for modeling of interactions with a crossbasis function.
This limitation meant that we were unable to model a potential size by environment (i.e. drought history) interaction, which may occur with many plant species [@tredennickSizebyenvironmentInteractionsNeglected2018].
To check for a size by environment interaction, we followed the methods of Tredennick et al. [-@tredennickSizebyenvironmentInteractionsNeglected2018] and fit a pair of generalized linear mixed effects models (GLMMs) for each vital rate using the *lme4* package [@batesFittingLinearMixedeffects2015].
Rather than explicitly modeling the effects of drought history, we modeled environmental stochasticity with random effects of year.
For each vital rate, we essentially replaced the crossbasis function in Eq 1.
with a random effect of year on either the intercept or both the intercept and relationship with plant size.
We assessed the significance of the year-by-size interaction by testing the likelihood ratio of models with and without the interaction against an expected likelihood ratio distribution under the null hypothesis of no interaction [@tredennickSizebyenvironmentInteractionsNeglected2018].

# Conclusion

# Acknowledgments

We thank ----, ----, ----, and --- anonymous reviewers for helpful discussions and comments on the manuscript.
We also thank the technicians and students who helped conduct the *Heliconia* censuses and the BDFFP and INPA for logistical support.
Financial support was provided by the U.S.
National Science Foundation (awards DEB-0614339, DEB-0614149, INT 98-06351, and DEB-1948607).
This is publication No. ---- in the BDFFP Technical Series.

If TRMM data is used, they suggest: "The TMPA data were provided by the\
NASA/Goddard Space Flight Center's Mesoscale Atmospheric Processes Laboratory and\
PPS, which develop and compute the TMPA as a contribution to TRMM.
"

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```