### 2021-03-16 README.Rmd
remove non-applicable info from README, now that this is no longer a package.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# HeliconiaDemography

<!-- badges: start -->

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/BrunaLab/HeliconiaDemography/master?urlpath=rstudio) <!-- [![R build status](https://github.com/BrunaLab/HeliconiaDemography/workflows/R-CMD-check/badge.svg)](https://github.com/BrunaLab/HeliconiaDemography/actions) -->

<!-- badges: end -->

This repository contains the data and code for our paper:

> Authors, (YYYY).
> *`r Title`*.
> Name of journal/book <https://doi.org/xxx/xxx>

<!-- Our pre-print is online here: -->

<!-- > Authors, (YYYY). *`r Title`*. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`).
> *Compendium of R code and data for `r Title`*.
> Accessed `r format(Sys.Date(), "%d %b %Y")`.
> Online at <https://doi.org/xxx/xxx>

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on [binder](https://mybinder.org/v2/gh/BrunaLab/HeliconiaDemography/master?urlpath=rstudio) to open an instance of RStudio in your browser, which will have the compendium files ready to work with.
Binder uses rocker-project.org Docker images to ensure a consistent and reproducible computational environment.
These Docker images can also be used locally.

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip).
After unzipping:

-   open the `.Rproj` file in RStudio
-   something about `renv`
-   something about `targets::tar_make()` and `targets::tar_visnetwork()` <!-- -   run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the [DESCRIPTION](/DESCRIPTION) file). --> <!-- -   finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console -->

### Licenses

<!-- **Text and figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/) -->

<!-- <!-- **Code :** See the [DESCRIPTION](DESCRIPTION) file -->

<!-- **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->