### 2021-05-13 doc/supplemental.Rmd
combine supplemental into one doc

---
title: "Supplement to 'Delayed effects of climate on vital rates leads to demographic divergence in Amazonian forest fragments'"
author: 
  - Eric R. Scott
  - María Uriarte
  - Emilio M. Bruna
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      linenumbers: no
      number_sections: no
      fig_caption: yes
      # reference_docx: "templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=templates/scholarly-metadata.lua
      - --lua-filter=templates/author-info-blocks.lua
      - --lua-filter=templates/pagebreak.lua
linkcolor: blue
bibliography: references.bib
csl: "templates/global-change-biology.csl" # Insert path for the bib-style
nocite: |
  @mcphadenChildProdigy1997981999, marengoDroughtAmazonia20052008, zengCausesImpacts20052008, lewis2010AmazonDrought2011
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

withr::with_dir(here::here(), {
  targets::tar_load(c(
    normals_plot,
    embrapa_wide,
    xa_spei,
    trmm_spei,
    gspei
    ))
})

library(tidyverse)
library(patchwork)
library(flextable)
library(tsibble)
library(lubridate)
library(latex2exp)
library(colorspace)
library(glue)
source(here::here("R", "utils.R"))
```

# BDFFP plot map

Map here

Map caption here

\pagebreak

# Climate Normals for BDFFP

```{r normals}
normals_plot & theme(text = element_text(size = 10))
```

Figure S1.
Weather normals for Manaus, Brazil (3º6'S, 60º1'W) for 1981--2010.
Precipitation (a) shows a marked dry seasons from June through October.
Temperature (b) varies less throughout the year.
Mean monthly temperature is shown in the solid red line and monthly minimum and maximum temperatures are shown with the lower and upper dashed lines, respectively.
Data from Brazilian National Institute of Meterology (INMET).
<!--# citation needed? -->

\pagebreak

# Comparison of SPEI calculated from different data sources

```{r}
# combine data
xa_df <-
  xa_spei %>% 
  as_tibble() %>% 
  group_by(yearmonth) %>% 
  #Summarize across grid cells
  summarize(xa_spei = mean(spei))

trmm_df <-
  trmm_spei %>% 
  group_by(yearmonth) %>% 
  summarize(trmm_spei = mean(trmm_spei))

comb <-
  full_join(xa_df, gspei) %>% 
  full_join(trmm_df) %>% 
  full_join(embrapa_wide %>% select(yearmonth, st_spei)) %>% 
  pivot_longer(c(-yearmonth), names_to = "dataset", values_to = "spei") %>% 
  filter(yearmonth > yearmonth(ymd("1990-01-01"))) %>% 
  mutate(year = year(yearmonth), .before = yearmonth)
```

```{r}
# data for plot
spei_plot <- 
  comb %>% 
  group_by(year) %>% 
  mutate(min_mon = spei == min(spei, na.rm = TRUE)) %>% 
  mutate(dataset = case_when(
    dataset == "g_spei"  ~ "Global SPEI",   
    dataset == "xa_spei" ~ "Xavier et al.",
    dataset == "st_spei" ~ "EMBRAPA",
    dataset == "trmm_spei" ~ "TRMM"
  ))
```

```{r spei-plot, fig.width=7, fig.height=3.5}

p <-
  ggplot(spei_plot, aes(x = yearmonth, y = spei)) +
  geom_line(aes(color = dataset), size = 0.5) +
  scale_x_yearmonth(
    "Date",
    date_breaks = "1 year",
  ) +
  scale_y_continuous(TeX("SPEI_3_"),
                     breaks = seq(-2.5, 2.5, 0.5))+
  scale_color_discrete_qualitative() +
  coord_cartesian(xlim = c("1995-02-01", "2009-02-01")) +
  labs(color = "Dataset", linetype = "Dataset") +
  theme_bw() +
  theme(text = element_text(size = 11), axis.text.x = element_text(angle = 45, hjust = 1))

annotate_spei(p)
```

Figure S2.
Three-month SPEI at BDFFP calculated using 4 datasets.
"EMBRAPA" (red) is the average of SPEI calculated using precipitation and evapotranspiration from two weather stations near BDFFP---Estacao Rio Preto da Eva ( -2.65º, -59.65º) and Estacao Manaus (-2.65º, -60.35º).
The data was accessed from the EMBRAPA Infoclima data portal (<https://www.cnpaf.embrapa.br/infoclima/>).
"Global SPEI" (green) is from the SPEI global database[@vicente-serranoNewGlobalGridded2010; @begueriaMultiscalarGlobalDrought2010].
"TRMM" (blue) SPEI uses precipitation data from TRMM [@trmmTMPA3B432011] and evapotranspiration data from @xavierDailyGriddedMeteorological2016 to calculate 3-month SPEI.
Note that TRMM data only goes back to 1998.
@xavierDailyGriddedMeteorological2016 is the SPEI source described in the main text.
Drought categories mild, moderate, severe, and extreme drought are represented by yellow, goldenrod, orange, and red panels, respectively.

\pagebreak

```{r}
citations <- tribble(~year, ~Reference,
                     2005, "Marengo et al. 2008; Zeng et al. 2008",
                     2010, "Lewis et al. 2011",
                     1997, "McPhaden 1999",
                     2002, NA,
                     2009, NA,
                     2004, NA,
                     2006, NA) 
```

```{r}
spei_table <-
  comb %>% 
  group_by(year, dataset) %>% 
  mutate(min = min(spei)) %>% 
  filter(spei == min) %>% 
  mutate(month = month(yearmonth, label = TRUE)) %>% 
  # summarize(min = min, month = month) %>%
  filter(year %in% c(2002, 2004, 2006, 2009, 1997, 2005, 2010)) %>% 
  select(year, dataset, spei, month) %>% 
  mutate(dataset = str_remove(dataset, "_spei")) %>% 
  pivot_wider(names_from = dataset, values_from = c(spei, month)) %>% 
  add_column(cause = c("ENSO", "ENSO", "ENSO", "AMO", "ENSO", "ENSO", "AMO"), 
             .after = "year") %>% 
  mutate(`Global SPEI` = glue("{round(spei_g, 2)} ({month_g})"),
         `Xavier et al.` = glue("{round(spei_xa, 2)} ({month_xa})"),
         EMBRAPA = glue("{round(spei_st, 2)} ({month_st})"),
         TRMM = glue("{round(spei_trmm, 2)} ({month_trmm})")) %>% 
  mutate(TRMM = if_else(TRMM == "NA (NA)", glue("–"), TRMM)) %>% 
  ungroup()

spei_table <- left_join(spei_table, citations)
```

```{r}
set_flextable_defaults(big.mark = "")
table <- 
  spei_table %>% 
  # select(-starts_with("month_"), -starts_with("spei_")) %>% 
  flextable(
    col_keys = c("year", "cause", "Global SPEI", "Xavier et al.", "EMBRAPA", "TRMM", "BDFFP")
  ) %>% 
  add_header_row(colwidths = c(1,1, 5),
                 values = c("", "", "Lowest SPEI for the year (month)")) %>% 
  align(i = 1, part = "header", align = "center") %>% 
  #extreme drought
  bg(~ spei_g < -2, ~`Global SPEI`, bg = "#e31a1c") %>% 
  bg(~ spei_xa < -2, ~ `Xavier et al.`, bg = "#e31a1c") %>% 
  bg(~ spei_st < -2, ~ EMBRAPA, bg = "#e31a1c") %>% 
  bg(~spei_trmm < -2, ~TRMM, bg = "#e31a1c") %>%

  #severe drought
  bg(~ between(spei_g,     -2, -1.5), ~ `Global SPEI`, bg = "#fd8d3c") %>% 
  bg(~ between(spei_xa,    -2, -1.5), ~ `Xavier et al.`, bg = "#fd8d3c") %>% 
  bg(~ between(spei_st,    -2, -1.5), ~ EMBRAPA, bg = "#fd8d3c") %>% 
  bg(~ between(spei_trmm,  -2, -1.5), ~ TRMM, bg = "#fd8d3c") %>%

  #moderate drought
  bg(~ between(spei_g,     -1.5, -1), ~ `Global SPEI`, bg = "#fecc5c") %>% 
  bg(~ between(spei_xa,    -1.5, -1), ~ `Xavier et al.`, bg = "#fecc5c") %>% 
  bg(~ between(spei_st,    -1.5, -1), ~ EMBRAPA, bg = "#fecc5c") %>% 
  bg(~ between(spei_trmm,  -1.5, -1), ~ TRMM, bg = "#fecc5c") %>%

  #mild drought
  bg(~ between(spei_g,     -1, 0), ~ `Global SPEI`, bg = "#ffffb2") %>% 
  bg(~ between(spei_xa,    -1, 0), ~ `Xavier et al.`, bg = "#ffffb2") %>% 
  bg(~ between(spei_st,    -1, 0), ~ EMBRAPA, bg = "#ffffb2") %>% 
  bg(~ between(spei_trmm,  -1, 0), ~ TRMM, bg = "#ffffb2") 
```

Table S1.
Comparison of SPEI values and drought categories across datasets for major drought events in the Amazon.
Mild, moderate, severe, and extreme drought are represented by yellow, goldenrod, orange, and red, respectively.
Cause indicates whether the drought was caused by the El Niño Southern Oscillation (ENSO) or Atlantic Sea Surface Temperatures (AMO).
The lowest SPEI value for the year identified by each dataset is reported, along with the month that minimum value occurred in.
Datasets are described in the legend for Figure S2.

```{r spei-table}
table %>%
  set_table_properties(layout = "autofit") %>%
  align(align = "center", j = 3:6) %>%
  border_remove() %>% 
  hline_bottom() %>% 
  hline_top() %>% 
  hline(j = 3:6, part = "header") %>% 
  footnote(j = 2, i = c(1, 4, 7), value = as_paragraph(
    c("McPhaden 1999",
      "Marengo et al. 2008; Zeng et al. 2008",
      "Lewis et al. 2011")),
    ref_symbols = c("a", "b", "c"),
    part = "body", inline = TRUE
  )
```

\pagebreak

# Choice of variable for plant size

include contents of `choose_size_var.Rmd` here.
Ask Emilio about citaiton for dataset.

# References

::: {#refs}
:::