### 2021-05-13 doc/validate_models.Rmd

---
title: "Model validation plots"
author: "Eric R. Scott"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## target knits Rmds in their own session, so load libraries here.
library(targets)
library(statmod)
library(tidyverse)
library(gratia)
```

```{r load-targets}
withr::with_dir(here::here(),{
  tar_load(c(s_cf, s_1ha, g_cf, g_1ha, f_cf, f_1ha))
  })
```

# Analysis

## Survival

Quantile residuals are more appropriate for binomial models.

### Continuous Forest

```{r}
summary(s_cf)
```

```{r}
draw(s_cf)
```

```{r echo=TRUE}
par(mfrow = c(2,2))
#histogram
plot(density(qresid(s_cf)))
#QQ plot
qqnorm(qresid(s_cf)); qqline(qresid(s_cf))
#Fitted vs. residuals
scatter.smooth(predict(s_cf, type = "response"), qresid(s_cf), col = "grey")
```

### Fragments

```{r}
summary(s_1ha)
```

```{r}
draw(s_1ha)
```

```{r echo=TRUE}
par(mfrow = c(2,2))
#histogram
plot(density(qresid(s_1ha)))
#QQ plot
qqnorm(qresid(s_1ha)); qqline(qresid(s_1ha))
#Fitted vs. residuals
scatter.smooth(predict(s_1ha, type = "response"), qresid(s_1ha), col = "grey")
```

## Growth

### Continuous Forest

```{r}
summary(g_cf)
```

```{r}
draw(g_cf)
```

```{r}
gratia::appraise(g_cf)
```

### Fragments

```{r}
summary(g_1ha)
```

```{r}
draw(g_1ha)
```

```{r}
appraise(g_1ha)
```

## Flowering

Quantile residuals are more appropriate for binomial models.

### Continuous Forest

```{r}
summary(f_cf)
```

```{r}
draw(f_cf)
```

```{r}
par(mfrow = c(2,2))
#histogram
plot(density(qresid(f_cf)))
#QQ plot
qqnorm(qresid(f_cf)); qqline(qresid(f_cf))
#Fitted vs. residuals
scatter.smooth(predict(f_cf, type = "response"), qresid(f_cf), col = "grey")
```

### Fragments

```{r}
summary(f_1ha)
```

```{r}
draw(f_1ha)
```

```{r}
par(mfrow = c(2,2))
#histogram
plot(density(qresid(f_1ha)))
#QQ plot
qqnorm(qresid(f_1ha)); qqline(qresid(f_1ha))
#Fitted vs. residuals
scatter.smooth(predict(f_1ha, type = "response"), qresid(f_1ha), col = "grey")
```

# Reproducibility

<details>

<summary>

Reproducibility receipt

</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>