### 2021-08-11 doc/validate_models.Rmd
re-organize validation

---
title: "Model validation plots"
author: "Eric R. Scott"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## target knits Rmds in their own session, so load libraries here.
library(targets)
library(statmod)
library(tidyverse)
library(gratia)
library(mgcv)
library(dlnm)
library(arm)
```

**TODO**

- increase knots for some models (https://github.com/BrunaLab/HeliconiaDemography/issues/92)
- check diagnostics for individual-level random effects.  Doesn't look good, but not sure how to interpret or what to do.

# Survival

```{r, echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(s_1ha, s_cf, s_cf_sub))
})
```

## Forest Fragments

### Diagnostics

```{r echo=TRUE}
s_1ha_res <- qresid(s_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(s_1ha_res))
#QQ plot
qqnorm(s_1ha_res); qqline(s_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_1ha), residuals(s_1ha, type = "response"))
```

### Basis dimensions

```{r}
# gam.check(s_1ha)
```

Basis dimension checking with `gam.check()` doesn't appear to work for `dlnm` crossbasis smooths. Instead I'll use a method described in the help file `?choose.k` to check for adequate knots.  Unfortunately, `bs = "cs"` doesn't work with `dlnm`, so I'll use `select = TRUE` instead to reduce chance of overfitting.

```{r}
# looking for near zero edf
gam(s_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_1ha)) #fine

gam(s_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_1ha)) #fine
```

### Summary

```{r}
summary(s_1ha)
```

```{r}
draw(s_1ha)
```

## Continuous Forest

### Diagnostics

```{r echo=TRUE}
s_cf_res <- qresid(s_cf)
par(mfrow = c(2,2))
#histogram
plot(density(s_cf_res))
#QQ plot
qqnorm(s_cf_res); qqline(s_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_cf), residuals(s_cf, type = "response"))
```
### Basis dimensions

```{r}
# looking for near zero edf
gam(s_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_cf)) #fine

gam(s_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_cf)) 
```

### Summary

```{r}
summary(s_cf)
```

```{r}
draw(s_cf)
```

## Effect of sample size

Here I use a random sub-sample to check that differences between continuous forest and fragments are not purely due to sample size differences, particularly differences in the complexity of the crossbasis smooth.

```{r}
summary(s_1ha)$edf[3]
summary(s_cf)$edf[3]
summary(s_cf_sub)$edf[3]
```

```{r}
draw(s_cf_sub)
```
Surface still looks different in a similar way though.

```{r echo=FALSE}
rm(s_1ha, s_cf, s_cf_sub)
```

# Growth

```{r echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(g_1ha, g_cf, g_cf_sub))
})
```

## Fragments

### Diagnostics

```{r}
appraise(g_1ha)
```

### Basis dimensions

```{r}
g_1ha_res <- residuals(g_1ha)
# looking for near zero edf
gam(g_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_1ha)) #fine

gam(g_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(g_1ha)) #fine
```

### Summary

```{r}
summary(g_1ha)
```

```{r}
draw(g_1ha)
```

## Continuous Forest

### Diagnostics

```{r}
gratia::appraise(g_cf)
```

### Basis dimensions

```{r}
g_cf_res <- residuals(g_cf)
# looking for near zero edf
gam(g_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_cf)) #fine

gam(g_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(g_cf)) #fine
```

### Summary

```{r}
summary(g_cf)
```

```{r}
draw(g_cf)
```

## Continuous Forest subsample

To check that differences are not purely due to sample size differences, particularly that lower edf in continuous forests is due to higher sample size.

```{r}
summary(g_1ha)$edf[3]
summary(g_cf)$edf[3]
summary(g_cf_sub)$edf[3]
```
edf of subsample is similar to full dataset, despite differences in sample size.

```{r}
draw(g_cf_sub)
```
Crossbasis surface looks nearly identical.

```{r echo=FALSE}
rm(g_1ha, g_cf, g_cf_sub)
```

# Flowering

```{r echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(f_1ha, f_cf, f_cf_sub))
})
```

## Fragments

### Diagnostics

```{r}
f_1ha_res <- qresid(f_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(f_1ha_res))
#QQ plot
qqnorm(f_1ha_res); qqline(f_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_1ha), residuals(f_1ha, type = "response"))
```

### Basis dimensions

```{r}
# looking for near zero edf
gam(f_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(f_1ha)) #fine

gam(f_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(f_1ha)) #fine
```

### Summary

```{r}
summary(f_1ha)
```

```{r}
draw(f_1ha)
```

## Continuous Forest

### Diagnostics

```{r}
f_cf_res <- qresid(f_cf)
par(mfrow = c(2,2))
#histogram
plot(density(f_cf_res))
#QQ plot
qqnorm(f_cf_res); qqline(f_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_cf), residuals(f_cf, type = "response"))
```

### Basis dimensions

```{r}
# looking for near zero edf
gam(f_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(f_cf)) #fine

gam(f_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(f_cf)) #fine
```

### Summary

```{r}
summary(f_cf)
```

```{r}
draw(f_cf)
```

## Continuous Forest subsample

To check that differences are not purely due to sample size differences, particularly that lower edf in continuous forests is due to higher sample size.  For flowering, edf is actually slightly higher in CF with a larger sample size.  With subsample, edf are more similar.

```{r}
summary(f_1ha)$edf[3]
summary(f_cf)$edf[3]
summary(f_cf_sub)$edf[3]
```

```{r}
draw(f_cf_sub)
```
Crossbasis surface looks extremely similar.

# Reproducibility

<details>

<summary>

Reproducibility receipt

</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>