### 2021-08-23 README.Rmd
closes #92 (but does not update results yet)

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# HeliconiaDemography

<!-- badges: start -->

[![Run Tests](https://github.com/BrunaLab/HeliconiaDemography/actions/workflows/run-testthat.yaml/badge.svg)](https://github.com/BrunaLab/HeliconiaDemography/actions/workflows/run-testthat.yaml)

<!-- badges: end -->

This repository contains the data and code for a manuscript written for submission to Global Change Biology tentatively titled: "Delayed effects of climate on vital rates leads to demographic divergence in Amazonian forest fragments"

## How to run

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip).

To run the compendium and reproduce all outputs:

-   Open the project in RStudio by double-clicking the `HeliconiaDemography.Rproj` file.
-   Install the [`renv` package](https://rstudio.github.io/renv/articles/renv.html) if it's not already installed and run `renv::restore()` to install all package dependencies.
-   Install the [`targets` package](https://docs.ropensci.org/targets/).
-   Run `targets::tar_make()` or `targets::tar_make_clustermq()` from the R console to run all code and produce all outputs.

**NOTE**: The full pipeline will take a *very long* time to run.  I suggest switching off individual-level random effects (more details in comments in `_targets.R`) if you want to run the project on your own computer.  Without individual-level random effects the pipeline should run relatively quickly (< 1hr, I think).

**Using HPC**:
Also, you may need to install packages on the remote HPC if using `tar_make_clustermq()` with the SSH option.  Unfortunately, I'm not sure if there is an easy way to do this using `renv`. You should, however, make sure the same versions of the packages are installed locally and on the HPC, especially `gratia` which I'm using a development version of that has important bugs fixed, but not new bugs introduced.

## Additional Dependencies

-   Numbering of and in-text references to equations in `paper.Rmd` is handled by [`pandoc-crossref`](https://github.com/lierdakil/pandoc-crossref), which may not be part of the standard RStudio pandoc installation. You might need to install it for the main text to knit correctly to Word.

<!--
### Licenses
-->
<!-- **Text and figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/) -->

<!-- <!-- **Code :** See the [DESCRIPTION](DESCRIPTION) file -->

<!-- **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->